---
title: "Classify case-control in kinetic state tremors using random forest models with median and iqr features, for Multiple Sclerosis. The training is done using controls from mpower and elevateMS, and cases from elevateMS. Gender removed as a factor variable in the Random forest model"
author: "Meghasyam Tummalacherla"
author options: "This would not have been possible without Thanneer Perumal's skeleton code"
date: "`r base::date()`"
output: html_document
editor_options: 
chunk_output_type: console
---
```{r knit2synapse, eval=FALSE}
library(synapser)
library(knit2synapse)

synapser::synLogin()

knit2synapse::createAndKnitToFolderEntity(file = "./mpower_control_elevateMS_no_gender.Rmd",
                                 parentId = "syn19120636",
                                 folderName = "elevateMS action tremor test (mpower PD action tremor train)(no scaling)(atleast 2 records)(no gender)")
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file
###########################################################
## Required libraries
###########################################################

library(CovariateAnalysis) # get it from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)

library(synapser)
library(knitr)
library(githubr)

library(ggplot2)
library(parsedate)
library(lubridate)
library(ranger)
library(ROCR)
library(Boruta)
library(MatchIt)
library(caret)
library(GA)

synapser::synLogin()

library(foreach)
library(doParallel)

cl = makeCluster(detectCores()-2)
registerDoParallel(cl)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE,
  cache.lazy = FALSE)
```

```{r github.synapse.parameters, include=FALSE, cache=TRUE}
###########################################################
## Github Parameters and links
###########################################################
# Synapse parameters
parentId = 'syn19120636';
activityName = 'elevateMS Kinetic tremor case-control classification';
activityDescription = 'Classify case-control in kinetic state tremors using random forest models with median and iqr features, for Multiple Sclerosis. The training is done using controls from mpower and elevateMS, and cases from elevateMS'

# Github link
gtToken = '~/github_token.txt'
githubr::setGithubToken(as.character(read.table(gtToken)$V1))

thisFileName <- 'mpower_control_elevateMS_no_gender.Rmd'

thisRepo <- getRepo(repository = "itismeghasyam/elevateMS_analysis", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('analysis/Tremor_analysis/no_gender/',thisFileName))
```

```{r fxn}
## Function block
## Function to classify case-control differences in each assay (balanced data)
rfModelCaseControl <- function(dat, responseNames, contFeatures, alwaysSplitVariables){
  # Get data and format data
  dat[,c(responseNames)] = lapply(dat[,c(responseNames), drop = F], factor)
  dat[,contFeatures] = lapply(dat[,contFeatures, drop = F], as.numeric)
  
  # Split training and testing data (based on healthCodes)
  hc = as.character(unique(dat$healthCode))
  hc.train = sample(hc, round(length(hc)*0.6))
  hc.test = setdiff(hc, hc.train)
  
  # Get training data
  dat.train = dat %>%
    dplyr::filter(healthCode %in% hc.train) %>%
    dplyr::select(one_of(contFeatures), 'MS') %>%
    dplyr::mutate(MS = as.numeric(MS) - 1)
  dat.train[is.na(dat.train)] = 0
  
  # Get testing data
  dat.test = dat %>%
    dplyr::filter(healthCode %in% hc.test) %>%
    dplyr::select(one_of(contFeatures), 'MS') %>%
    dplyr::mutate(MS = as.numeric(MS) - 1)
  dat.test[is.na(dat.test)] = 0

  tryCatch({
    # Fit a rf model
    rf.mdl = ranger(dependent.variable.name = responseNames, 
                    data = dat.train, 
                    num.trees = 1000,
                    importance = 'permutation',
                    seed = 123456,
                    classification = TRUE,
                    write.forest = TRUE)
  
    # Predict reponse of test data
    rf.prediction = predict(rf.mdl, 
                            data = dat.test,
                            predict.all = T,
                            type = 'response',
                            seed = 123456)
  
    pred.probabilities = rowSums(rf.prediction$predictions == 1, na.rm = T)/rf.prediction$num.trees
  
    mdl.prediction = pred.probabilities %>%
      prediction(predictions = ., labels = dat.test[,'MS'], label.ordering = c(0,1))
    
    auroc = data.frame(metricName = 'auroc', 
                       value = performance(mdl.prediction, 'auc')@y.values[[1]])
    acc = data.frame(metricName = 'acc',
                     value = max(performance(mdl.prediction, 'acc')@y.values[[1]]))
    
    rec = performance(mdl.prediction, 'rec')@y.values[[1]]
    pre = performance(mdl.prediction, 'prec')@y.values[[1]]
    
    aupr = data.frame(metricName = 'aupr', 
                      value = caTools::trapz(rec[!is.nan(rec) & !is.nan(pre)], 
                                             pre[!is.nan(rec) & !is.nan(pre)]))
    mdl.performance = rbind(auroc, acc, aupr)
  
    mdl.coeff = ranger::importance(rf.mdl) %>%
      rownameToFirstColumn('featureNames') %>%
      dplyr::rename(value = DF)
  
    mdl.param = rbind(
      data.frame(metricName = 'nData', value = dim(dat)[1]),
      data.frame(metricName = 'nTrain', value = dim(dat.train)[1]),
      data.frame(metricName = 'nTest', value = dim(dat.test)[1]),
      data.frame(metricName = 'nTrainControl', value = sum(dat.train[,'MS'] == 0)),
      data.frame(metricName = 'nTestControl', value = sum(dat.test[,'MS'] == 1))
    )
    
    all.results = list(param = mdl.param, performance = mdl.performance, coeff = mdl.coeff)
  }, error = function(e){
    all.results = list(param = NA, performance = NA, coeff = NA)
  })
}

## Function to display number of data points
printNumData <- function(mdls){
  mapply(function(x, y){
    x[[1]]$param %>% 
      plyr::rename(c('value'=y))
  }, mdls, names(mdls), SIMPLIFY = F) %>% 
    plyr::join_all()
}

## Function to plot and display average model performance
printModelPerformance <- function(mdl, assayName){
  tmp = purrr::imap(mdl, function(x, y){
    names(x) = paste0('perm.',1:length(x))
    tmp = purrr::imap(x, function(y,z){
      
      datTempInner <- y$performance
      if(is.na(datTempInner)){
        datTempInner <- data.frame(c('auroc','acc','aupr'),rep(NA,3))
        colnames(datTempInner) <- c('metricName','value')
      }
      return(datTempInner)
    }) %>%
      plyr::join_all(by = 'metricName')}) 
  
  ##### DATA RESHAPE IS NEEDED
  tmpa <- data.frame(rep(NA,3*(length(tmp$alternateModel)-1)),rep(NA,3*(length(tmp$alternateModel)-1)),rep(NA,3*(length(tmp$alternateModel)-1)))
  colnames(tmpa) <- c('metricName','value','modelName')
  tmpa$metricName <- rep(tmp$alternateModel$metricName,length(tmp$alternateModel)-1)
  a <- unlist(tmp$alternateModel)
  tmpa$value <- a[4:(3*length(tmp$alternateModel))]
  tmpa$modelName <- rep('alternateModel', length(tmpa$value))
  
  tmpb <- data.frame(rep(NA,3*(length(tmp$alternateModel)-1)),rep(NA,3*(length(tmp$alternateModel)-1)),rep(NA,3*(length(tmp$alternateModel)-1)))
  colnames(tmpb) <- c('metricName','value','modelName')
  tmpb$metricName <- rep(tmp$nullModel$metricName,length(tmp$alternateModel)-1)
  a <- unlist(tmp$nullModel)
  tmpb$value <- a[4:(3*length(tmp$nullModel))]
  tmpb$modelName <- rep('nullModel', length(tmpb$value))
  
  tmp <- rbind(tmpa,tmpb)
  
  p = ggplot(tmp, aes(x = metricName, y = value)) + geom_boxplot(aes(color = modelName), position = position_identity())
  p = p + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylim(0,1) + ggtitle(assayName) + scale_color_manual(values = c('black','grey60'))
  
  tmp = dplyr::group_by(tmp, modelName, metricName) %>%
    dplyr::summarise(value = round(mean(value, na.rm = T),4)) %>%
    tidyr::spread(modelName, value) 
  return(list(plot = p, performance = tmp))
}

## Function to print top n features 
printTopFeatures <- function(mdl){
  purrr::imap(mdl, function(x, i){
    x$coeff %>%
      plyr::rename(c('value' = i))
    }) %>%
    plyr::join_all(by = 'featureNames') %>%
    tidyr::gather(iter, value, -featureNames) %>%
    dplyr::group_by(featureNames) %>%
    dplyr::summarise(value = sum(value != 0, na.rm = T)/100) %>%
    dplyr::arrange(desc(value)) %>%
    dplyr::filter(value >= 0.9) %>%
    dplyr::select(featureNames) %>%
    unlist %>% 
    paste(collapse = ', ')
}

## Get top n percent of features 
getTopNPercentOfFtrs <- function(mdl, n = 70){
  purrr::imap(mdl, function(x, i){
    x$coeff %>%
      plyr::rename(c('value' = i))
  }) %>%
    plyr::join_all() %>%
    tidyr::gather(iter, value, -featureNames) %>%
    dplyr::group_by(featureNames) %>%
    dplyr::summarise(value = sum(value != 0, na.rm = T)/100) %>%
    dplyr::arrange(desc(value)) %>%
    dplyr::filter(value >= n/100) %>%
    dplyr::select(featureNames) %>%
    unlist %>% 
    setdiff('(Intercept)')
}
```

#### Download age matched elevateMS (control + cases) + mpower (cases) pool
```{r age.matched.healthcodes, fig.height=5, fig.width=10}
###########################################################
## Download data
###########################################################
age.records.matched.id = 'syn19123754'
age.records.matched.syn = synapser::synGet(age.records.matched.id)
age.records.matched = read.csv(age.records.matched.syn$path, sep = '\t')
all.used.ids = age.records.matched.id

p1 = ggplot(age.records.matched, aes(x = factor(MS), y = age)) + geom_boxplot()
p2 = ggplot(age.records.matched, aes(x = MS)) + geom_bar()
p3 = ggplot(age.records.matched, aes(x = log10(n), color = factor(MS))) + geom_density() + xlab('log10(# Windows)')
ggpubr::ggarrange(p2,p1, p3,ncol = 3, nrow = 1)
```

```{r get.summarized.ftrs}

###########################################################
## Download and set up features for elevateMS
###########################################################
summarized.ftrs.elevateMS.id = 'syn18879962'
summarized.ftrs.elevateMS = read.csv(synapser::synGet(summarized.ftrs.elevateMS.id)$path,
                                     sep = '\t')
all.used.ids = c(all.used.ids, summarized.ftrs.elevateMS.id)
###########################################################
## Download demohgraphics data for elevateMS
###########################################################
# Get demographics from synapse
demo.tbl.id.ms = 'syn17115631' # elevateMS_baselineCharacteristics
demo.tbl.syn.ms <- synapser::synGet(demo.tbl.id.ms)
demo.tbl.ms <- demo.tbl.syn.ms$path %>% read.csv(sep = '\t')
all.used.ids <- c(all.used.ids, demo.tbl.id.ms)

###########################################################
## Download and set up features for age matched mpower controls
###########################################################
summarized.ftrs.mpower.id = 'syn19165094'
summarized.ftrs.mpower = read.csv(synapser::synGet(summarized.ftrs.mpower.id)$path,
                                     sep = '\t')
all.used.ids <- c(all.used.ids, summarized.ftrs.mpower.id)

```

### Features considered for this analysis
```{r ftr.def, echo = TRUE}
# Split features based componenet
ftrs.names = c("mean.tm", "median.tm", "mode.tm", "mx.tm", "mn.tm", "sd.tm", "skewness.tm", 
               "kurtosis.tm", "Q25.tm", "Q75.tm", "range.tm", "rough.tm", "rugo.tm", "energy.tm",
               "mobility.tm", "mtkeo.tm", "dfa.tm", "rmsmag.tm", "IQR.tm", "complexity.tm",
               "mn.fr", "mx.fr", "sd.fr", "sem.fr", "md.fr", "mod.fr", "Q25.fr", "Q75.fr", 
               "IQR.fr", "cent.fr", "skew.fr", "kurt.fr", "sfm.fr", "sh.fr", "ewt.permEnt.fr",
               "ewt.shannonEnt.fr", "ewt.simpsonEnt.fr", "ewt.renyiEnt.fr", "ewt.tsallisEnt.fr", 
               "EnergyInBand1", "EnergyInBand1_5", "EnergyInBand2", "EnergyInBand2_5", "EnergyInBand3",
               "EnergyInBand3_5", "EnergyInBand4", "EnergyInBand4_5", "EnergyInBand5", "EnergyInBand5_5",
               "EnergyInBand6", "EnergyInBand6_5", "EnergyInBand7", "EnergyInBand7_5", "EnergyInBand8",
               "EnergyInBand8_5", "EnergyInBand9", "EnergyInBand9_5", "EnergyInBand10", "EnergyInBand10_5",
               "EnergyInBand11", "EnergyInBand11_5", "EnergyInBand12", "EnergyInBand12_5", "EnergyInBand13",
               "EnergyInBand13_5", "EnergyInBand14", "EnergyInBand14_5", "EnergyInBand15", "EnergyInBand15_5",
               "EnergyInBand16", "EnergyInBand16_5", "EnergyInBand17", "EnergyInBand17_5", "EnergyInBand18",
               "EnergyInBand18_5", "EnergyInBand19", "EnergyInBand19_5", "EnergyInBand20", "EnergyInBand20_5",
               "EnergyInBand21", "EnergyInBand21_5", "EnergyInBand22", "EnergyInBand22_5", "EnergyInBand23",
               "EnergyInBand23_5", "EnergyInBand24", "EnergyInBand24_5")
```


```{r filter.important.features}

ranking.syn.id = 'syn12038526' # AllRanking.Rdata (feature ranking  postural, kinetic and rest for mpower)
load(synapser::synGet(ranking.syn.id)$path)
all.used.ids <- c(all.used.ids, ranking.syn.id)
nFeatures = 100 # max number of features to choose per assay

kineticFeaturesIMF1 <- ranka[,3] %>% na.omit() 
kineticFeaturesIMF2 <- ranka[,4] %>% na.omit() 
maxRank <- max(c(kineticFeaturesIMF1,kineticFeaturesIMF2))
kineticFeaturesIMF1 <- kineticFeaturesIMF1[kineticFeaturesIMF1 > (maxRank-nFeatures)] %>% names()
kineticFeaturesIMF2 <- kineticFeaturesIMF2[kineticFeaturesIMF2 > (maxRank-nFeatures)] %>% names()

kineticFeaturesToSelect <- c()
for (ftr in colnames(summarized.ftrs.elevateMS)){
  ftr1 <- gsub('IMF1.','',ftr)
  # print(ftr1)
  if(ftr1 %in% kineticFeaturesIMF1){
    kineticFeaturesToSelect <- c(kineticFeaturesToSelect,ftr)
  }
  ftr1 <- gsub('IMF2.','',ftr)
  # print(ftr1)
  if(ftr1 %in% kineticFeaturesIMF2){
    kineticFeaturesToSelect <- c(kineticFeaturesToSelect,ftr)
  }
}

# For training data add both the ftrs from elevateMS and the controls from 
# mpower and filter to get the whole age matched data
dfMainTrain <- dplyr::full_join(summarized.ftrs.elevateMS, summarized.ftrs.mpower) %>% 
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>%
  dplyr::select(c('healthCode',kineticFeaturesToSelect)) %>% 
  unique()

```
#### Compute model performance using identified features (from mpoweranalysis)
##### Evaluate model on training set (mpower + elevateMS) (alternate vs null model)
```{r all.rf.model}
nperm = 100
set.seed(123456789)
exportPackages = c('dplyr','plyr','data.table','CovariateAnalysis', 'ranger', 'ROCR')
ftrs.to.evaluate = kineticFeaturesToSelect %>% unique()

# Get training data
dat.train = dfMainTrain %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
  dplyr::select(one_of(ftrs.to.evaluate)) 
# %>%
  # apply(2, scale) 
dat.train = cbind(data.frame(healthCode = dfMainTrain %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
    dplyr::select(healthCode)), dat.train) %>%
  dplyr::left_join(age.records.matched %>%
                     dplyr::select(healthCode,MS,age,gender)) %>% 
  as.data.frame() %>% droplevels() %>%
  unique()

# Run random forest model
rf.mdl.caret =  c('alternateModel','nullModel') %>%
  plyr::llply(function(mdl.tag, datTrain, nperm, ftrsNames, rfModelCaseControl){
    mdl = plyr::llply(seq(nperm), .fun = function(iperm, mdlTag, datTrain, ftrsNames, rfModelCaseControl){
      # Randomise case control label per healthCode
      if(mdlTag == 'nullModel')
        datTrain$MS = sample(datTrain$MS) 
      
      rfModelCaseControl(datTrain,  
                         responseNames = 'MS',
                         contFeatures = ftrsNames)
    }, mdl.tag, datTrain, ftrsNames, rfModelCaseControl,
    .parallel = F, 
    .paropts = list(.packages = exportPackages, 
                    .export = c('rfModelCaseControl')))
    return(mdl)
  },
  dat.train, nperm, setdiff(ftrs.to.evaluate, c('age', 'gender')), rfModelCaseControl,
  .paropts = list(.packages = exportPackages, 
                  .export = c('rfModelCaseControl')))
names(rf.mdl.caret) = c('alternateModel','nullModel')
```

##### Model performance
```{r all.print.results, fig.height=5, fig.width=12, cache = FALSE}
# Display number of data points used
writeLines('Number of data points used')
kable(rf.mdl.caret$alternateModel[[1]]$param)

# Plot auroc and aupr
writeLines('Average performance on 100 runs')
tmp = printModelPerformance(rf.mdl.caret, 'Caret based features')
p = gridExtra::tableGrob(tmp$performance, rows = NULL)
p = ggpubr::ggarrange(tmp$plot, p, nrow = 1, ncol = 2)
p

# Plot important variables in all model
writeLines('Features present in 70% of the models are')
printTopFeatures(rf.mdl.caret$alternateModel) %>%
  paste(collapse = ', ') %>%
  data.frame(Important_Features = .) %>%
  kable()
```

#### Evaluate Kinetic Model on elevateMS by training on mPower+elevateMS age matched data
```{r model.result}
nperm = 100
set.seed(123456789)
exportPackages = c('dplyr','plyr','data.table','CovariateAnalysis', 'ranger', 'ROCR')

# ftrsToSelect <- paste0(kineticFeaturesToSelect,'.kinetic')
ftrsToSelect <- kineticFeaturesToSelect
  
# Get training data
dat.train = dfMainTrain %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
  dplyr::select(one_of(ftrsToSelect)) 
# %>%
#   apply(2, scale) 
dat.train = cbind(data.frame(healthCode = dfMainTrain %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
    dplyr::select(healthCode)), dat.train) %>%
  dplyr::left_join(age.records.matched %>%
                     dplyr::select(healthCode,MS,age,gender)) %>% 
  as.data.frame() %>% droplevels() %>%
  unique()

dat.test <- summarized.ftrs.elevateMS %>%
  dplyr::filter(!(healthCode %in% age.records.matched$healthCode)) %>% 
    dplyr::select(healthCode,one_of(kineticFeaturesToSelect)) %>% 
  dplyr::left_join(demo.tbl.ms %>%
                     dplyr::select(healthCode, gender, age, MS = dataGroups)) %>%
  dplyr::filter(MS %in% c('ms_patient','control'),
                gender %in% c('male','female'),
                !is.na(age), !is.na(MS), !is.na(gender)) %>% #remove NAs
  dplyr::mutate(MS = factor(MS, levels = c('control','ms_patient')),
                MS = as.numeric(MS) - 1) %>% 
  unique() %>% 
  droplevels()

    rf.mdl = ranger(dependent.variable.name = 'MS', 
                    data = dat.train, 
                    num.trees = 1000,
                    importance = 'permutation',
                    seed = 123456,
                    classification = TRUE,
                    write.forest = TRUE)
  
    # Predict reponse of test data
    rf.prediction = predict(rf.mdl, 
                            data = dat.test,
                            predict.all = T,
                            type = 'response',
                            seed = 123456)
        pred.probabilities = rowSums(rf.prediction$predictions == 1, na.rm = T)/rf.prediction$num.trees
  
    mdl.prediction = pred.probabilities %>%
      prediction(predictions = ., labels = dat.test[,'MS'], label.ordering = c(0,1))
    
    auroc = data.frame(metricName = 'auroc', 
                       value = performance(mdl.prediction, 'auc')@y.values[[1]])
    acc = data.frame(metricName = 'acc',
                     value = max(performance(mdl.prediction, 'acc')@y.values[[1]]))
    
    rec = performance(mdl.prediction, 'rec')@y.values[[1]]
    pre = performance(mdl.prediction, 'prec')@y.values[[1]]
    
    aupr = data.frame(metricName = 'aupr', 
                      value = caTools::trapz(rec[!is.nan(rec) & !is.nan(pre)], 
                                             pre[!is.nan(rec) & !is.nan(pre)]))
    mdl.performance = rbind(auroc, acc, aupr)
    
writeLines(c('Number of Kinetic features chosen for analysis : ',as.character(length(kineticFeaturesToSelect))))
writeLines(c('Number of healthCodes in testset : ',as.character(length(dat.test$healthCode))))
writeLines('Prediction on elevateMS after training on age matched mPower + elevateMS data')
kable(mdl.performance)
```

### Store results in synapse
```{r syn.store, cache=FALSE, eval = TRUE}
CODE = Folder(name = 'elevateMS action tremor test (mpower PD action tremor train)(no scaling)(atleast 2 records)(no gender)', 
              parentId = "syn19120636")
CODE = synStore(CODE)

# Save all coefficients and AUCs as RData
save(list = c('rf.mdl'), 
     file = 'AllImportanceScore.RData')
obj = File('AllImportanceScore.RData', 
           name = 'All importance scores with AUC', 
           parentId = CODE$properties$id)
obj = synStore(obj,activityName = activityName, 
               activityDescription = activityDescription, 
               used = all.used.ids, executed = thisFile)
res = paste0('https://www.synapse.org/#!Synapse:',obj$properties$id)
# stopCluster(cl)
```
[Results](`r res`)

### Source code in github
[Source Code](`r thisFile`)