---
title: "elevateMS baseline characteristics vs objective tests"
author: "Abhi Pratap @ Sage Bionetworks"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r global_options, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
Sys.setenv(TZ='GMT')
library(install.load)
install_load(c('plyr', 'tidyverse', 'data.table', 'lubridate', 'ggpubr', 'synapser'))
install_load('ggthemes', 'devtools', 'gridExtra')
install_load("ComplexHeatmap", "circlize", "RColorBrewer")
synapser::synLogin()

###Load data freeze
load(synGet("syn19273733")$path)
source("~/dev/elevateMS_analysis/analysis/Analytical_Paper_2/common_analysis_functions.R")
setwd("~/dev/elevateMS_analysis/analysis/Analytical_Paper_2/")
```


```{r}
baselineChar_tmp <- baselineChar  %>%
  dplyr::filter(dataGroups == 'ms_patient') %>%
  dplyr::mutate(age_group = cut(as.numeric(age), breaks=c(17, 29, 39, 49, 59, 120)),
                age_group = revalue(age_group, replace=c('(17,29]' = '18-29','(29,39]' = '30-39',
                                                         '(39,49]' = '40-49','(49,59]' = '50-59',
                                                         '(59,120]' = '60+')),
                YearsSinceDiagnosis.grouped = cut(as.numeric(YearsSinceDiagnosis), breaks=c(1,4,10,18,100)),
                YearsSinceFirstDMT.grouped = cut(as.numeric(YearsSinceFirstDMT), breaks=c(1,4,10,18,100)),
                race.mod = ifelse(race == 'caucasian', 'Caucasian', 'Non-Caucasian'),
                race.mod = factor(race.mod),
                education = factor(education),
                gender = factor(gender),
                health_insurance.mod = ifelse(health_insurance == 'no insurance', NA, health_insurance),
                health_insurance.mod = factor(health_insurance.mod),
                msFamilyHistory = factor(msFamilyHistory),
                employment = factor(employment)) %>%
  dplyr::select(healthCode, overallPhysicalAbility, race.mod, age_group, education, employment,
                gender, health_insurance.mod, YearsSinceFirstDMT.grouped, 
                YearsSinceDiagnosis.grouped, msFamilyHistory ) %>%
  dplyr::mutate(overallPhysicalAbility = factor(overallPhysicalAbility, 
                                                levels=c('normal', 'milddisability', 
                                                         'moderatedisability', 'gaitdisability')))
```


-----


### Correlation between PDDS Survey(Overall Physical Disability) and other meta data 

```{r, fig.align='center', fig.width=8, fig.height=6}
library(heatmaply)
#devtools::install_github("th1vairam/CovariateAnalysis")
corOutput = CovariateAnalysis::getAssociationStatistics(baselineChar_tmp %>% 
                                              select(overallPhysicalAbility, health_insurance.mod,
                                                     YearsSinceDiagnosis.grouped, employment,
                                                     YearsSinceFirstDMT.grouped,msFamilyHistory,
                                                     age_group, education, race.mod))

ggsave("Figs_N_Tables/BaselinePDDS_vs_baselineChars_Cor.jpeg", plot = corOutput$plot, width = 8, height = 6, dpi = 200)
ggsave("Figs_N_Tables/BaselinePDDS_vs_baselineChars_Cor.tiff", plot = corOutput$plot, width = 8, height = 6, dpi = 200)


cors <- as.data.frame( corOutput$ESTIMATE) %>% 
  rownames_to_column(var='var1') %>% 
  tidyr::gather('var2', 'cor', -var1)

pvals <- as.data.frame( corOutput$PVAL) %>% 
  rownames_to_column(var='var1') %>% 
  tidyr::gather('var2', 'pval', -var1) 

res = merge(pvals, cors) %>% 
  mutate(pval = round(pval, digits = 4),
         cor = round(cor, digits=4)) %>%
  filter(var1 == 'overallPhysicalAbility') %>%
  select(-var1) %>%
  dplyr::arrange(desc(cor)) %>%
  rename(baselineChar = var2)

library(DT)
datatable(res, rownames = F)
```


------


### Association between Tap Features(summarized over time) and baseline characteristics survey
_We adjust for hand in the model_

```{r}
tmp_calc_kurtosis <- function(x, na.rm){
  e1071::kurtosis(x, na.rm = na.rm, type=1)
}

tapF_averaged <- tapF %>% 
  dplyr::filter(dataGroups == 'ms_patient') %>% 
  tidyr::gather(feature, value, c(TAP_FEATURES)) %>% 
  select(healthCode, hand, feature, value) %>%
  group_by(healthCode, hand, feature) %>% 
  dplyr::summarise_all(.funs=c(median=median, IQR=IQR), na.rm=T) %>%
  tidyr::gather(stat, value, c('median', 'IQR')) %>%
  dplyr::mutate(feature = paste0(feature,'.',stat)) %>% 
  dplyr::select(-stat)

baselineChars_vs_tapF <- baselineChar_tmp %>% 
  dplyr::select(healthCode, overallPhysicalAbility, YearsSinceDiagnosis.grouped,
                YearsSinceFirstDMT.grouped, msFamilyHistory,
                age_group, education, race.mod) %>%
  tidyr::gather(baselineChar, baselineValue, c(2:8)) %>%
  inner_join(tapF_averaged)


tmp_lm_model <- function(df){
  if(unique(df$baselineChar_tmp) == "overallPhysicalAbility"){
      df <- df %>% 
        dplyr::mutate(baselineValue = factor(baselineValue, 
                                             levels=c('normal', 'milddisability', 
                                                      'moderatedisability', 'gaitdisability')))
    }
    lm(value ~ baselineValue + hand, data=df) 
}

association_baselineChars_vs_tapF <- baselineChars_vs_tapF  %>%
  dplyr::mutate(baselineChar_tmp = baselineChar) %>%
  group_by(feature, baselineChar) %>%
  nest() %>%
  mutate(lm_res = purrr::map(data, tmp_lm_model),
         anova_res = purrr::map(lm_res, function(x){ anova(x) }) ) 

lm.model.effectSize <- association_baselineChars_vs_tapF %>%
  mutate(glance = lm_res %>% map(broom::tidy)) %>%
  select(-data, -lm_res, -anova_res) %>%
  unnest() %>%
  dplyr::filter(str_detect(term, "baselineValue")) %>%
  dplyr::select(baselineChar, feature, term, lmModel.effectSize = estimate, lmModel.pVal = p.value)

anova.pvals <- association_baselineChars_vs_tapF %>%
  mutate(glance = anova_res %>% map(broom::tidy)) %>%
  select(-data, -lm_res, -anova_res) %>%
  unnest() %>% 
  dplyr::filter(term == 'baselineValue') %>%
  group_by(baselineChar) %>%
  nest() %>%
  mutate(res = map(data, function(df) { data.frame(feature = df$feature,p.value = df$p.value,
                                                   p.val.adj = p.adjust(df$p.value, method="fdr") ) } )) %>%
  select(-data) %>% 
    unnest() %>%
  rename(anova.p.val = p.value,
         anova.p.val.adj = p.val.adj)

##Feature with best association with PDDS
bestTapF_PDDS_feature  <- anova.pvals %>%
  dplyr::filter(baselineChar == 'overallPhysicalAbility') %>%
  dplyr::slice(which.min(anova.p.val.adj)) %>% .$feature %>% as.character()

bestTapF_vs_baselineChar <- anova.pvals %>%
  dplyr::filter(feature == bestTapF_PDDS_feature) %>%
  dplyr::select(baselineChar, feature, anova.p.val.adj) %>%
  dplyr::mutate(anova.p.val.adj = round(anova.p.val.adj, digits = 5))

datatable(lm.model.effectSize %>% full_join(anova.pvals) %>%
            dplyr::select(baselineChar, feature, term, lmModel.effectSize, lmModel.pVal, anova.p.val, anova.p.val.adj),
          extensions = 'Buttons', 
          options = list(dom = 'Bfrtip', buttons = c('csv', 'excel')),
          rownames = F)
```

<br><br>


#### Median number of taps vs top baseline characteristics

```{r, fig.align='center', fig.width=8, fig.height=6}
#1 
df <- baselineChars_vs_tapF %>%
  dplyr::filter(baselineChar == "overallPhysicalAbility", feature == 'numberTaps.median') %>%
  filter(!is.na(baselineValue)) %>%
  dplyr::mutate(baselineValue = gsub('disability', '-disability', baselineValue),
                  baselineValue = factor(baselineValue, 
                                         levels=c('normal', 'mild-disability',
                                                  'moderate-disability', 'gait-disability')))

p <- ggplot(data = df, aes(y=value, x=(baselineValue))) + geom_boxplot(width=0.5) + geom_jitter(size=0.3, width=0.2, color="grey60") + theme_bw(base_size = 15)
p <- p + xlab('Physical ability category') + ylab("Number of median taps (Finger tapping)") + coord_flip()
p
ggsave("Figs_N_Tables/BaselinePDDS_vs_TapF1.jpeg", plot = p, width = 8, height = 6, dpi = 200)
ggsave("Figs_N_Tables/BaselinePDDS_vs_TapF1.tiff", plot = p, width = 8, height = 6, dpi = 200)


#2 
df <- baselineChars_vs_tapF %>%
  dplyr::filter(baselineChar == "age_group", feature == 'numberTaps.median') %>%
  filter(!is.na(baselineValue))
p <- ggplot(data = df, aes(y=value, x=(baselineValue))) + geom_boxplot(width=0.5) + geom_jitter(size=0.3, width=0.2, color="grey60") + theme_bw(base_size = 15) 
p <- p + xlab('Age group') + ylab("Number of median taps(Finger tapping)") + coord_flip()
p
ggsave("Figs_N_Tables/BaselineAge_vs_TapF1.jpeg", plot = p, width = 8, height = 6, dpi = 200)
ggsave("Figs_N_Tables/BaselineAge_vs_TapF1.tiff", plot = p, width = 8, height = 6, dpi = 200)

```


<br><br>

--------


### Association between Walk Features(summarized over time) and baseline characteristics survey

```{r}
walkF_averaged <- walkF %>% 
  dplyr::filter(dataGroups == 'ms_patient') %>% 
  tidyr::gather(feature, value, c(WALK_FEATURES)) %>% 
  select(healthCode, feature, value) %>%
  group_by(healthCode, feature) %>% 
  dplyr::summarise_all(.funs=c(median=median, IQR=IQR), na.rm=T) %>%
  tidyr::gather(stat, value, c('median', 'IQR')) %>%
  dplyr::mutate(feature = paste0(feature,'.',stat)) %>% 
  dplyr::select(-stat)

baselineChars_vs_walkF <- baselineChar_tmp %>% 
   dplyr::select(healthCode, overallPhysicalAbility, YearsSinceDiagnosis.grouped,
                YearsSinceFirstDMT.grouped, msFamilyHistory,
                age_group, education, race.mod) %>%
  tidyr::gather(baselineChar, baselineValue, c(2:8)) %>%
  inner_join(walkF_averaged)


tmp_lm_model <- function(df){
  if(unique(df$baselineChar_tmp) == "overallPhysicalAbility"){
      df <- df %>% 
        dplyr::mutate(baselineValue = factor(baselineValue, 
                                             levels=c('normal', 'milddisability', 
                                                      'moderatedisability', 'gaitdisability')))
    }
    lm(value ~ baselineValue, data=df) 
}

association_baselineChars_vs_walkF <- baselineChars_vs_walkF  %>%
  mutate(baselineChar_tmp = baselineChar) %>%
  group_by(feature, baselineChar) %>%
  nest() %>%
  mutate(lm_res = purrr::map(data, tmp_lm_model),
         anova_res = purrr::map(lm_res, function(x){ anova(x) }) ) 

lm.model.effectSize <- association_baselineChars_vs_walkF %>%
  mutate(glance = lm_res %>% map(broom::tidy)) %>%
  select(-data, -lm_res, -anova_res) %>%
  unnest() %>%
  dplyr::filter(str_detect(term, "baselineValue")) %>%
  dplyr::select(baselineChar, feature, term, lmModel.effectSize = estimate, lmModel.pVal = p.value)

anova.pvals <- association_baselineChars_vs_walkF %>%
  mutate(glance = anova_res %>% map(broom::tidy)) %>%
  select(-data, -lm_res, -anova_res) %>%
  unnest() %>% 
  dplyr::filter(term == 'baselineValue') %>%
  group_by(baselineChar) %>%
  nest() %>%
  mutate(res = map(data, function(df) { data.frame(feature = df$feature,p.value = df$p.value,
                                                   p.val.adj = p.adjust(df$p.value, method="fdr") ) } )) %>%
  select(-data) %>% 
    unnest() %>%
  rename(anova.p.val = p.value,
         anova.p.val.adj = p.val.adj)
  

##Feature with best association with PDDS
bestWalkF_PDDS_feature  <- anova.pvals %>%
  dplyr::filter(baselineChar == 'overallPhysicalAbility') %>%
  dplyr::slice(which.min(anova.p.val.adj)) %>% .$feature

bestWalkF_vs_baselineChar <- anova.pvals %>% 
  dplyr::filter(feature == bestWalkF_PDDS_feature) %>%
  dplyr::select(baselineChar, feature, anova.p.val.adj) %>%
  dplyr::mutate(anova.p.val.adj = round(anova.p.val.adj, digits = 5))

datatable(lm.model.effectSize %>% full_join(anova.pvals) %>% dplyr::arrange(anova.p.val.adj) %>%
            dplyr::select(baselineChar, feature, term, lmModel.effectSize, lmModel.pVal, anova.p.val, anova.p.val.adj), 
          extensions = 'Buttons', 
          options = list(dom = 'Bfrtip', buttons = c('csv', 'excel')),
          rownames = F)
```



```{r, fig.align='center', fig.width=8, fig.height=6}
df <- baselineChars_vs_walkF %>%
  dplyr::filter(baselineChar == "overallPhysicalAbility", feature == 'medianF0FAJ.median') %>%
  filter(!is.na(baselineValue)) %>%
  dplyr::mutate(baselineValue = gsub('disability', '-disability', baselineValue),
                  baselineValue = factor(baselineValue, 
                                         levels=c('normal', 'mild-disability',
                                                  'moderate-disability', 'gait-disability')))

p <- ggplot(data = df,
       aes(y=log10(value), x=baselineValue)) + geom_boxplot(width=0.5) + geom_jitter(size=0.3, width=0.2, color="grey60") +  theme_bw()  + theme_bw(base_size = 15) + xlab('Physical ability category') + ylab("log10(Average jerk derived feature (walking))")  + coord_flip()
p
ggsave("Figs_N_Tables/BaselinePDDS_vs_WalkF1.jpeg", plot = p, width = 8, height = 6, dpi = 200)
ggsave("Figs_N_Tables/BaselinePDDS_vs_WalkF1.tiff", plot = p, width = 8, height = 6, dpi = 200)
```


---------

<br>


### Association between Tremor Features(summarized over time) and baseline characteristics survey


```{r}
tremorF_averaged <- tremorF %>% 
  dplyr::filter(dataGroups == 'ms_patient') %>% 
  tidyr::gather(feature, value, c(TREMOR_FEATURES)) %>% 
  select(healthCode, feature, value, hand) %>%
  group_by(healthCode, feature, hand) %>% 
  dplyr::summarise_all(.funs=c(median=median, IQR=IQR), na.rm=T) %>%
  tidyr::gather(stat, value, c('median', 'IQR')) %>% as.data.frame() %>%
  dplyr::mutate(feature = paste0(feature,'.',stat)) %>% 
  dplyr::select(-stat)

baselineChars_vs_tremorF <- baselineChar_tmp %>% 
   dplyr::select(healthCode, overallPhysicalAbility, YearsSinceDiagnosis.grouped,
                YearsSinceFirstDMT.grouped, msFamilyHistory,
                age_group, education, race.mod) %>%
  tidyr::gather(baselineChar, baselineValue, c(2:8)) %>%
  inner_join(tremorF_averaged)


tmp_lm_model <- function(df){
  if(unique(df$baselineChar_tmp) == "overallPhysicalAbility"){
      df <- df %>% 
        dplyr::mutate(baselineValue = factor(baselineValue, 
                                             levels=c('normal', 'milddisability', 
                                                      'moderatedisability', 'gaitdisability')))
    }
    lm(value ~ baselineValue + hand, data=df) 
}

association_baselineChars_vs_tremorF <- baselineChars_vs_tremorF  %>%
  dplyr::mutate(baselineChar_tmp = baselineChar) %>%
  group_by(feature, baselineChar) %>%
  nest() %>%
  mutate(lm_res = purrr::map(data, tmp_lm_model),
         anova_res = purrr::map(lm_res, function(x){ anova(x) }) ) 

lm.model.effectSize <- association_baselineChars_vs_tremorF %>%
  mutate(glance = lm_res %>% map(broom::tidy)) %>%
  select(-data, -lm_res, -anova_res) %>%
  unnest() %>%
  dplyr::filter(str_detect(term, "baselineValue")) %>%
  dplyr::select(baselineChar, feature, term, lmModel.effectSize = estimate, lmModel.pVal = p.value) %>%
  dplyr::filter(!is.na(lmModel.pVal))

anova.pvals <- association_baselineChars_vs_tremorF %>%
  mutate(glance = anova_res %>% map(broom::tidy)) %>%
  select(-data, -lm_res, -anova_res) %>%
  unnest() %>% 
  dplyr::filter(term == 'baselineValue') %>%
  group_by(baselineChar) %>%
  nest() %>%
  mutate(res = map(data, function(df) { data.frame(feature = df$feature,p.value = df$p.value,
                                                   p.val.adj = p.adjust(df$p.value, method="fdr") ) } )) %>%
  select(-data) %>% 
  unnest() %>%
  dplyr::filter(!is.na(p.value)) %>%
  rename(anova.p.val = p.value,
         anova.p.val.adj = p.val.adj)

##Feature with best association with PDDS
bestTremorF_PDDS_feature  <- anova.pvals %>%
  dplyr::filter(baselineChar == 'overallPhysicalAbility') %>%
  dplyr::slice(which.min(anova.p.val.adj)) %>% .$feature %>% as.character()

bestTremorF_vs_baselineChar <- anova.pvals %>% 
  dplyr::filter(feature == bestTremorF_PDDS_feature) %>%
  dplyr::select(baselineChar, feature, anova.p.val.adj) %>%
  dplyr::mutate(anova.p.val.adj = round(anova.p.val.adj, digits = 5))

datatable(anova.pvals %>% inner_join(lm.model.effectSize) %>% dplyr::arrange(anova.p.val.adj) %>%
            dplyr::select(baselineChar, feature, term, lmModel.effectSize, lmModel.pVal, anova.p.val, anova.p.val.adj), 
          extensions = 'Buttons', 
          options = list(dom = 'Bfrtip', buttons = c('csv', 'excel')),
          rownames = F)
```



```{r, fig.align='center', fig.width=8, fig.height=6}
df <- baselineChars_vs_tremorF %>%
  dplyr::filter(baselineChar == "overallPhysicalAbility", 
                feature == 'skewness.tm.IMF1.md_uav_gyroscope.IQR') %>%
  filter(!is.na(baselineValue)) %>%
  dplyr::mutate(baselineValue = gsub('disability', '-disability', baselineValue),
                  baselineValue = factor(baselineValue, 
                                         levels=c('normal', 'mild-disability',
                                                  'moderate-disability', 'gait-disability')))
p <- ggplot(data = df,
       aes(y=value, x=baselineValue)) + geom_boxplot(width=0.5) + geom_jitter(size=0.3, width=0.2, color="grey60") +  theme_bw()  + theme_bw(base_size = 15) 
p <- p + xlab('Physical ability category') + ylab("Hand rotational velocity derived feature (finger-to-nose)")  + coord_flip()
p
ggsave("Figs_N_Tables/BaselinePDDS_vs_TremorF1.jpeg", plot = p, width = 8, height = 6, dpi = 200)
ggsave("Figs_N_Tables/BaselinePDDS_vs_TremorF1.tiff", plot = p, width = 8, height = 6, dpi = 200)
```

----

<br>

### Association between Rest Features(summarized over time) and baseline characteristics survey

```{r}
restF_averaged <- restF %>% 
  dplyr::filter(dataGroups == 'ms_patient') %>% 
  tidyr::gather(feature, value, c(REST_FEATURES)) %>% 
  select(healthCode, feature, value) %>%
  group_by(healthCode, feature) %>% 
  dplyr::summarise_all(.funs=c(median=median, IQR=IQR), na.rm=T) %>%
  tidyr::gather(stat, value, c('median', 'IQR')) %>%
  dplyr::mutate(feature = paste0(feature,'.',stat)) %>% 
  dplyr::select(-stat)

baselineChars_vs_restF <- baselineChar_tmp %>% 
   dplyr::select(healthCode, overallPhysicalAbility, YearsSinceDiagnosis.grouped,
                YearsSinceFirstDMT.grouped, msFamilyHistory,
                age_group, education, race.mod) %>%
  tidyr::gather(baselineChar, baselineValue, c(2:8)) %>%
  inner_join(restF_averaged)

association_baselineChars_vs_restF <- baselineChars_vs_restF  %>%
  dplyr::mutate(baselineChar_tmp = baselineChar) %>%
  group_by(feature, baselineChar) %>%
  nest() %>%
  mutate(lm_res = purrr::map(data, function(x) lm(value ~ baselineValue, data=x)),
         anova_res = purrr::map(lm_res, function(x){ anova(x) }) ) 

lm.model.effectSize <- association_baselineChars_vs_restF %>%
  mutate(glance = lm_res %>% map(broom::tidy)) %>%
  select(-data, -lm_res, -anova_res) %>%
  unnest() %>%
  dplyr::filter(str_detect(term, "baselineValue")) %>%
  dplyr::select(baselineChar, feature, term, lmModel.effectSize = estimate, lmModel.pVal = p.value) %>%
  dplyr::filter(!is.na(lmModel.pVal))

anova.pvals <- association_baselineChars_vs_restF %>%
  mutate(glance = anova_res %>% map(broom::tidy)) %>%
  select(-data, -lm_res, -anova_res) %>%
  unnest() %>% 
  dplyr::filter(term == 'baselineValue') %>%
  group_by(baselineChar) %>%
  nest() %>%
  mutate(res = map(data, function(df) { data.frame(feature = df$feature,p.value = df$p.value,
                                                   p.val.adj = p.adjust(df$p.value, method="fdr") ) } )) %>%
  select(-data) %>% 
  unnest() %>%
  dplyr::filter(!is.na(p.value)) %>%
  rename(anova.p.val = p.value,
         anova.p.val.adj = p.val.adj)

##Feature with best association with PDDS
bestRestF_PDDS_feature  <- anova.pvals %>%
  dplyr::filter(baselineChar == 'overallPhysicalAbility') %>%
  dplyr::slice(which.min(anova.p.val.adj)) %>% .$feature %>% as.character()

bestRestF_vs_baselineChar <- anova.pvals %>% 
  dplyr::filter(feature == bestRestF_PDDS_feature) %>%
  dplyr::select(baselineChar, feature, anova.p.val.adj) %>%
  dplyr::mutate(anova.p.val.adj = round(anova.p.val.adj, digits = 5))


datatable(anova.pvals %>% inner_join(lm.model.effectSize) %>% 
            dplyr::arrange(anova.p.val.adj) %>%
            dplyr::select(baselineChar, feature, term, lmModel.effectSize, lmModel.pVal, anova.p.val, anova.p.val.adj), 
          extensions = 'Buttons', 
          options = list(dom = 'Bfrtip', buttons = c('csv', 'excel')),
          rownames = F)
```


---------

### Association between DSST Features (summarized over time) and baseline characteristics survey


```{r}
DSST_FEATURES <- c("numDigits", "numCorrect", "percentCorrect", 
                   "avgTime", "sdTime", "totalTime")

dsstF_averaged <- dsst %>% 
  dplyr::filter(dataGroups == 'ms_patient') %>% 
  tidyr::gather(feature, value, c(DSST_FEATURES)) %>% 
  select(healthCode, feature, value) %>%
  group_by(healthCode, feature) %>% 
  dplyr::summarise_all(.funs=c(median=median, IQR=IQR), na.rm=T) %>%
  tidyr::gather(stat, value, c('median', 'IQR')) %>%
  dplyr::mutate(feature = paste0(feature,'.',stat)) %>% 
  dplyr::select(-stat)

baselineChars_vs_dsstF <- baselineChar_tmp %>% 
   dplyr::select(healthCode, overallPhysicalAbility, YearsSinceDiagnosis.grouped,
                YearsSinceFirstDMT.grouped, msFamilyHistory,
                age_group, education, race.mod) %>%
  tidyr::gather(baselineChar, baselineValue, c(2:8)) %>%
  inner_join(dsstF_averaged)



tmp_lm_model <- function(df){
  if(unique(df$baselineChar_tmp) == "overallPhysicalAbility"){
      df <- df %>% 
        dplyr::mutate(baselineValue = factor(baselineValue, 
                                             levels=c('normal', 'milddisability', 
                                                      'moderatedisability', 'gaitdisability')))
    }
    lm(value ~ baselineValue, data=df) 
}

association_baselineChars_vs_dsstF <- baselineChars_vs_dsstF  %>%
  mutate(baselineChar_tmp = baselineChar) %>%
  group_by(feature, baselineChar) %>%
  nest() %>%
  mutate(lm_res = purrr::map(data, tmp_lm_model),
         anova_res = purrr::map(lm_res, function(x){ anova(x) }) ) 


lm.model.effectSize <- association_baselineChars_vs_dsstF %>%
  mutate(glance = lm_res %>% map(broom::tidy)) %>%
  select(-data, -lm_res, -anova_res) %>%
  unnest() %>%
  dplyr::filter(str_detect(term, "baselineValue")) %>%
  dplyr::select(baselineChar, feature, term, lmModel.effectSize = estimate, lmModel.pVal = p.value) %>%
  dplyr::filter(!is.na(lmModel.pVal))

anova.pvals <- association_baselineChars_vs_dsstF %>%
  mutate(glance = anova_res %>% map(broom::tidy)) %>%
  select(-data, -lm_res, -anova_res) %>%
  unnest() %>% 
  dplyr::filter(term == 'baselineValue') %>%
  group_by(baselineChar) %>%
  nest() %>%
  mutate(res = map(data, function(df) { data.frame(feature = df$feature,p.value = df$p.value,
                                                   p.val.adj = p.adjust(df$p.value, method="fdr") ) } )) %>%
  select(-data) %>% 
    unnest() %>%
  rename(anova.p.val = p.value,
         anova.p.val.adj = p.val.adj)

##Feature with best association with PDDS
bestDSSTF_PDDS_feature  <- anova.pvals %>%
  dplyr::filter(baselineChar == 'overallPhysicalAbility') %>%
  dplyr::slice(which.min(anova.p.val.adj)) %>% .$feature %>% as.character()

bestDSSTF_vs_baselineChar <- anova.pvals %>% 
  dplyr::filter(feature == bestDSSTF_PDDS_feature) %>%
  dplyr::select(baselineChar, feature, anova.p.val.adj) %>%
  dplyr::mutate(anova.p.val.adj = round(anova.p.val.adj, digits = 5))

datatable(anova.pvals %>% inner_join(lm.model.effectSize) %>% 
            dplyr::arrange(anova.p.val.adj) %>%
            dplyr::select(baselineChar, feature, term, lmModel.effectSize, lmModel.pVal, anova.p.val, anova.p.val.adj), 
          extensions = 'Buttons', 
          options = list(dom = 'Bfrtip', buttons = c('csv', 'excel')),
          rownames = F)


```



```{r, fig.align='center', fig.width=8, fig.height=6}
df <- baselineChars_vs_dsstF %>%
  dplyr::filter(baselineChar == "overallPhysicalAbility", feature == 'numCorrect.median') %>%
  filter(!is.na(baselineValue)) %>%
  dplyr::mutate(baselineValue = gsub('disability', '-disability', baselineValue),
                  baselineValue = factor(baselineValue, 
                                         levels=c('normal', 'mild-disability',
                                                  'moderate-disability', 'gait-disability')))
p <- ggplot(data = df,
       aes(y=(value), x=baselineValue)) + geom_boxplot(width=0.5) + geom_jitter(size=0.3, width=0.2, color="grey60") +  theme_bw()  + theme_bw(base_size = 15) + xlab('Physical ability category') + ylab("Number correctly called symbols (DSST)")  + coord_flip()
p
ggsave("Figs_N_Tables/BaselinePDDS_vs_DSSTF1.jpeg", plot = p, width = 8, height = 6, dpi = 200)
ggsave("Figs_N_Tables/BaselinePDDS_vs_DSSTF1.tiff", plot = p, width =8, height = 6, dpi = 200)
```


```{r, syn.store, eval=F}
# Knit html document and push to synapse
thisFileDir = '~/dev/elevateMS_analysis/analysis/Analytical_Paper_2/'
setwd(thisFileDir)
thisFilePrefix = "BaselineChars_vs_ObjectiveTasks"
thisFileName = paste0(thisFilePrefix, '.Rmd')
html.obj = rmarkdown::render(thisFileName, output_format = 'html_document')

thisFilePrefix = "BaselineChars_vs_ObjectiveTasks"
thisFileName = paste0(thisFilePrefix, '.Rmd')
gtToken = '~/.ssh/apratap_github_token_20190219.txt';
githubr::setGithubToken(as.character(read.table(gtToken)$V1))
thisRepo <- githubr::getRepo(repository = "Sage-Bionetworks/elevateMS_analysis", ref="branch", refName='master')
thisFile_github_permaLink <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('analysis/Analytical_Paper_2/',thisFileName))

##Internal Project
thisFile_synapse = File(paste0(thisFilePrefix, '.html'), parentId ='syn11315047')
synStore(thisFile_synapse, executed=thisFile_github_permaLink)

##Public Project
thisFile_synapse = File(paste0(thisFilePrefix, '.html'), parentId ='syn22005477')
synStore(thisFile_synapse, executed=thisFile_github_permaLink)

unlink(paste0(thisFilePrefix, '.html'))
```



