---
title: "ElevateMS Daily Check-in's vs Weather"
author: "Abhi Pratap"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

Given that most participants' self-reported response to daily checkin PROs have very few responses in the "horrible" category, the following analysis assess the sensitivity of association between weather and daily checkin after removing the "horrible" responses 


```{r global_options, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
rm(list=ls())
Sys.setenv(TZ='GMT')
library(install.load)
install_load(c('plyr', 'tidyverse', 'data.table', 'lubridate', 'ggpubr', 'synapser'))
install_load('ggthemes', 'devtools')
# if (!requireNamespace("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# BiocManager::install("ComplexHeatmap", version = "3.8")
install_load("ComplexHeatmap", "circlize", "RColorBrewer")
synapser::synLogin()

###Load data freeze
load(synGet("syn19273733")$path)
setwd("~/dev/elevateMS_analysis/analysis/Analytical_Paper_2/")
source("common_analysis_functions.R")
```



```{r}
dailyCheckins_flt <- dailyCheckins %>% 
  dplyr::filter(dataGroups == 'ms_patient') %>%
  dplyr::mutate(checkin_date = as.Date(activity_start_timestamp_GMT )) %>%
  dplyr::select(healthCode, checkin_date, pain=mood_pain, 
                mobility=mood_mobility, health=mood_health,
                participant_day, participant_week) %>%
  ## The following mapping obtained from elevateMS bridge project thru Larsson
  dplyr::mutate(pain = recode(pain, '1'='None', '2'='Mild',  '3'='Moderate', '4'='Severe' , '5'='Horrible'),
                pain = factor(pain, levels=c('None', 'Mild', 'Moderate', 'Severe' ,'Horrible')),
                mobility = recode(mobility, '1'='Excellent', '2'='Very Good', '3'='Good', '4'='Not Great', '5'='Horrible'),
                mobility = factor(mobility, levels=c('Excellent', 'Very Good', 'Good', 'Not Great', 'Horrible')),
                health = recode(health, '1'='Amazing', '2'='Okay', '3'='So-So', '4'='Not Great', '5'='Horrible'),
                health = factor(health, levels=c('Amazing', 'Okay', 'So-So', 'Not Great', 'Horrible')))

cat('pain \n')
prop.table(table(dailyCheckins_flt$pain)) * 100

cat('\n health \n')
prop.table(table(dailyCheckins_flt$health)) * 100

cat('\n mobility \n')
prop.table(table(dailyCheckins_flt$mobility)) * 100

# View(dailyCheckins_flt)
# table(dailyCheckins_flt$health, dailyCheckins_flt$mobility)

weatherF_flt <- weatherF %>%
  dplyr::mutate(weather_date = as.Date(timestamp)) %>%
  dplyr::select(healthCode, weather_date, atmosphericPressure, 
                cloudCoverage, currentTemperature, humidity, 
                maximumTemperature, minimumTemperature)


dailyCheckins_N_weather <- merge(dailyCheckins_flt, weatherF_flt, 
                            by.x=c('healthCode', 'checkin_date'), 
                            by.y=c('healthCode', 'weather_date')) %>%
  dplyr::mutate(pain = as.factor(pain),
                mobility = as.factor(mobility),
                health = as.factor(health))
```


### Daily Health vs Weather
_Baseline health level - Amazing_
```{r}
#### Mixed effect modelling
health <- dailyCheckins_N_weather %>% 
  filter(health != 'Horrible') %>% 
  mutate(health = as.character(health)) %>%
  mutate( health = factor(health, levels=c('Amazing', 'Okay', 'So-So', 'Not Great') ))
install_load("lme4", "lmerTest", "broom")
res.1 <- lmer(atmosphericPressure ~ health + (1 | healthCode), data=health )
res.2 <- lmer(cloudCoverage ~ health  + (1 | healthCode), data=health )
res.3 <- lmer(currentTemperature ~ health + (1 | healthCode), data=health )
res.4 <- lmer(humidity ~ health + (1 | healthCode), data=health )
res.5 <- lmer(maximumTemperature ~ health + (1 | healthCode), data=health )
res.6 <- lmer(minimumTemperature ~ health + (1 | healthCode), data=health )
sjPlot::tab_model(res.1, res.2, res.3, res.4, res.5, res.6, show.ci = F)
```


### Visual difference between weather features and daily health
```{r fig.align='center', fig.width=12, fig.height=4, dpi=100}
### Plot top differentiator - cloudCoverage
p1 <- ggplot(data=health, aes(x=health, y=currentTemperature)) + geom_boxplot(width=0.6)
p1 <- p1 + xlab('daily health score') + ylab("Temperature(F)")
p1 <- p1 + geom_jitter(size=0.3, width=0.2, color="grey60") + theme_bw(base_size = 15)

p2 <- ggplot(data=health, aes(x=health, y=humidity)) + geom_boxplot(width=0.6)
p2 <- p2 + xlab('daily health score') + ylab("Humidity")
p2 <- p2 + geom_jitter(size=0.3, width=0.2, color="grey60") + theme_bw(base_size = 15)

p3 <- ggplot(data=health, aes(x=health, y=maximumTemperature)) + geom_boxplot(width=0.6)
p3 <- p3 + xlab('daily health score') + ylab("Humidity")
p3 <- p3 + geom_jitter(size=0.3, width=0.2, color="grey60") + theme_bw(base_size = 15)

gridExtra::grid.arrange(p1, p2, p3, nrow=1)
```




-------


### Daily Mobility vs Weather
_Baseline Mobility level - Excellent_

```{r}
#### Mixed effect modelling
install_load("lme4", "lmerTest", "broom")
mobility <- dailyCheckins_N_weather %>% 
  filter(mobility != 'Horrible') %>% 
  mutate(mobility = as.character(mobility), 
         mobility = factor(mobility, levels=c('Excellent', 'Very Good', 'Good', 'Not Great')))

mobility.1 <- lmer(atmosphericPressure ~ mobility + (1 | healthCode), data=mobility )
mobility.2 <- lmer(cloudCoverage ~ mobility  + (1 | healthCode),  data=mobility )
mobility.3 <- lmer(currentTemperature ~ mobility + (1 | healthCode), data=mobility )
mobility.4 <- lmer(humidity ~ mobility + (1 | healthCode), data=mobility )
mobility.5 <- lmer(maximumTemperature ~ mobility + (1 | healthCode), data=mobility )
mobility.6 <- lmer(minimumTemperature ~ mobility + (1 | healthCode), data=mobility )
sjPlot::tab_model(mobility.1, mobility.2, mobility.3, mobility.4, mobility.5, mobility.6, show.ci = F)
```


### Visual difference between weather features and daily mobility
```{r fig.align='center', fig.width=6, fig.height=4, dpi=100}
p1 <- ggplot(data=mobility, aes(x=mobility, y=currentTemperature)) + geom_boxplot(width=0.6)
p1 <- p1 + xlab('daily mobility') +  ylab("Temperature(F)")
p1 <- p1 + geom_jitter(size=0.3, width=0.2, color="grey60") + theme_light(base_size = 15)
p1
```


-------

### Daily Pain vs Weather
_Baseline Pain level - None_


```{r}
#### Mixed effect modelling
install_load("lme4", "lmerTest", "broom")
pain <- dailyCheckins_N_weather %>% filter(pain != 'Horrible') %>% 
  mutate(pain = as.character(pain),
         pain = factor(pain, levels=c('None', 'Mild', 'Moderate', 'Severe')))
res.1 <- lmer(atmosphericPressure ~ pain + (1 | healthCode), data=pain )
res.2 <- lmer(cloudCoverage ~ pain  + (1 | healthCode),  data=pain )
res.3 <- lmer(currentTemperature ~ pain + (1 | healthCode), data=pain )
res.4 <- lmer(humidity ~ pain + (1 | healthCode), data=pain )
res.5 <- lmer(maximumTemperature ~ pain + (1 | healthCode), data=pain )
res.6 <- lmer(minimumTemperature ~ pain + (1 | healthCode), data=pain )
sjPlot::tab_model(res.1, res.2, res.3, res.4, res.5, res.6, show.ci = F)
```


```{r fig.align='center', fig.width=6, fig.height=4, dpi=100, eval=F}
### Visual difference between weather features and daily pain
### Plot top differentiator - cloudCoverage
p1  <- ggplot(data=pain, aes(x=pain, y=humidity)) + geom_boxplot(width=0.6)
p1  <- p1 + geom_jitter(size=0.3, width=0.2, color="grey60") + theme_light(base_size = 15)
p1  <- p1 + xlab('self reported daily pain')
p1
```



```{r, syn.store, eval=F}

# Knit html document and push to synapse
thisFilePrefix = "dailyCheckins_vs_Weather_SensitivityAnalysis"
thisFileName = paste0(thisFilePrefix, '.Rmd')
html.obj = rmarkdown::render(thisFileName, output_format = 'html_document')

thisFilePrefix = "dailyCheckins_vs_Weather_SensitivityAnalysis"
thisFileName = paste0(thisFilePrefix, '.Rmd')
gtToken = '~/.ssh/apratap_github_token_20190219.txt'
githubr::setGithubToken(as.character(read.table(gtToken)$V1))
thisRepo <- githubr::getRepo(repository = "Sage-Bionetworks/elevateMS_analysis", ref="branch", refName='master')
thisFile_github_permaLink <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('analysis/Analytical_Paper_2/',thisFileName))


###Store on Internal Synapse
thisFile_synapse = File(paste0(thisFilePrefix, '.html'), parentId ='syn11315047')
synStore(thisFile_synapse, executed=thisFile_github_permaLink)

###Store on Production Synapse
thisFile_synapse = File(paste0(thisFilePrefix, '.html'), parentId ='syn22005477')
synStore(thisFile_synapse, executed=thisFile_github_permaLink)

unlink(paste0(thisFilePrefix, '.html'))
```
