---
title: "elevateMS self-repored relpases"
author: "Abhi Pratap"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

----

```{r global_options, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
Sys.setenv(TZ='GMT')
library(install.load)
install_load(c('plyr', 'tidyverse', 'data.table', 'lubridate', 'ggpubr', 'synapser'))
install_load('ggthemes', 'devtools', "synapser")
synapser::synLogin()

###Load data freeze
load(synGet("syn19273733")$path)
```



```{r, echo=FALSE}
relapses <- relapses %>% 
  dplyr::mutate(relapse.date = as.Date(activity_start_timestamp_local))
relapses.flt <- relapses %>% filter(dataGroups == 'ms_patient')
```


1. How many controls report relapses: `r nrow(relapses %>% filter(dataGroups == 'control'))`
2. Cases reporting relapse: `r n_distinct(relapses.flt$healthCode)`
3. Number of relapses reported by MS patients(median): `r median(table(relapses.flt$healthCode))`

<br>
<br>


### Days when participants report relapse
```{r, fig.width=15, fig.height=12}
p <- ggplot(data=relapses.flt, aes(y=healthCode, x=participant_day)) + geom_point()
p <- p + theme_light(base_size = 15) + theme(axis.text.y = element_blank(),
                                             axis.ticks.y = element_blank())
p
```


<br>
<br>


### Number of times each app-based task is completed by participants before and after they report a relapse
```{r, fig.width=15, fig.height=12}
### Avg number of activities before and after all relapses
tmp_calcReplaseSummary <- function(userHealthCode, relapseDate){
  df <- userActivity %>% filter(healthCode == userHealthCode)
  before = df %>% filter(createdOn_localTime < relapseDate) %>% mutate(status = 'before')
  after = df %>% filter(createdOn_localTime > relapseDate) %>% mutate(status = 'after')
  res <- rbind.fill(before %>% select(status, originalTable), 
        after %>% select(status, originalTable)) %>%
    count(status, originalTable)
}
relapseSummary <- relapses.flt %>% 
  dplyr::group_by(healthCode, relapse.date) %>%
  do(relapseSummary = tmp_calcReplaseSummary(.$healthCode, .$relapse.date))
relapseSummary <- relapseSummary %>% unnest() %>%
  mutate(originalTable = gsub('-v.*', '', originalTable, perl = T))
p2 <- ggplot(data=relapseSummary, aes(x=status, y=log2(n)))  + geom_boxplot() + facet_wrap(. ~ originalTable)
p2 <- p2 + theme_minimal(base_size = 15)
p2
```




```{r, syn.store, eval=F}
# Knit html document and push to synapse
thisFileDir = '/Users/apratap/Dropbox/dev/elevateMS_analysis/analysis/Analytical_Paper_2/'
setwd(thisFileDir)
thisFilePrefix = "relapseAnalysis"
thisFileName = paste0(thisFilePrefix, '.Rmd')
html.obj = rmarkdown::render(thisFileName, output_format = 'html_document')

thisFilePrefix = "relapseAnalysis"
thisFileName = paste0(thisFilePrefix, '.Rmd')
gtToken = '~/.ssh/apratap_github_token_20190219.txt';
githubr::setGithubToken(as.character(read.table(gtToken)$V1))
thisRepo <- githubr::getRepo(repository = "Sage-Bionetworks/elevateMS_analysis", ref="branch", refName='master')
thisFile_github_permaLink <- githubr::getPermlink(repository = thisRepo, repositoryPath=paste0('analysis/Analytical_Paper_2/',thisFileName))
thisFile_synapse = File(paste0(thisFilePrefix, '.html'), parentId ='syn11315047')
###Store on Synapse
synStore(thisFile_synapse, executed=thisFile_github_permaLink)
unlink(paste0(thisFilePrefix, '.html'))
```
