---
title: "Classify case-control in kinetic state tremors using random forest models with median and iqr features, for Multiple Sclerosis. The training is done using controls from mpower and elevateMS, and cases from elevateMS"
author: "Meghasyam Tummalacherla"
author options: "This would not have been possible without Thanneer Perumal's skeleton code"
date: "`r base::date()`"
output: html_document
editor_options: 
chunk_output_type: console
---
```{r knit2synapse, eval=FALSE}
library(synapser)
library(knit2synapse)

synapser::synLogin()

knit2synapse::createAndKnitToFolderEntity(file = "./mpower_control_elevateMS.Rmd",
                                 parentId = "syn11315047",
                                 folderName = "elevateMS action tremor test (mpower PD action tremor train)")
```

```{r libs, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
## It is assumed your working directory is where this file
###########################################################
## Required libraries
###########################################################

library(CovariateAnalysis) # get it from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)

library(synapser)
library(knitr)
library(githubr)

library(ggplot2)
library(parsedate)
library(lubridate)
library(ranger)
library(ROCR)
library(Boruta)
library(MatchIt)
library(caret)
library(GA)

synapser::synLogin()

library(foreach)
library(doParallel)

cl = makeCluster(detectCores()-2)
registerDoParallel(cl)

options(xtable.type="html")

knitr::opts_chunk$set(
  echo=FALSE,
  warning=FALSE,
  message=FALSE,
  error = FALSE,
  tidy = FALSE,
  cache = TRUE,
  cache.lazy = FALSE)
```

```{r github.synapse.parameters, include=FALSE, cache=TRUE}
###########################################################
## Github Parameters and links
###########################################################
# Synapse parameters
parentId = 'syn11315047';
activityName = 'elevateMS Kinetic tremor case-control classification';
activityDescription = 'Classify case-control in kinetic state tremors using random forest models with median and iqr features, for Multiple Sclerosis. The training is done using controls from mpower and elevateMS, and cases from elevateMS'

# Github link
gtToken = '~/github_token.txt'
githubr::setGithubToken(as.character(read.table(gtToken)$V1))

thisFileName <- 'mpower_control_elevateMS.Rmd'

thisRepo <- getRepo(repository = "itismeghasyam/elevateMS_analysis", ref="branch", refName='master')
thisFile <- getPermlink(repository = thisRepo, repositoryPath=paste0('analysis/',thisFileName))
```

#### Age match elevateMS (control + cases) + mpower (cases) pool

```{r filter.healthcodes.elevateMS, fig.height=5, fig.width=10}
## Get and format data for elevateMS
### Age match case and control healthCodes
###########################################################
## Download data
###########################################################
# Tremor activity tables from elevateMS project
tremor.tbl.id = 'syn10278767' # Tremor Activity-v5
tremor.tbl.syn <- synapser::synTableQuery(paste0("SELECT * FROM ", tremor.tbl.id))
tremor.tbl <- tremor.tbl.syn$asDataFrame()

all.used.ids = tremor.tbl.id

# Get demographics from synapse
demo.tbl.id = 'syn10295288' # Demographics table-v2
demo.tbl.syn <- synapser::synTableQuery(paste0("SELECT * FROM ", demo.tbl.id))
demo.tbl <- demo.tbl.syn$asDataFrame()
all.used.ids <- c(all.used.ids, demo.tbl.id)

# Get Profile data (age data)
profile.tbl.id <- 'syn10235463'
profile.tbl.syn <- synapser::synTableQuery(paste0("SELECT * FROM ", profile.tbl.id))
profile.tbl <- profile.tbl.syn$asDataFrame() 
all.used.ids <- c(all.used.ids, profile.tbl.id)

# Certain healthCodes have multiple ages 
# (62, 63 etc., so we just consider the minimum of those two)
profile.tbl.cleaned <- profile.tbl %>% 
  dplyr::select(healthCode, age = demographics.age) %>% 
  dplyr::group_by(healthCode) %>% 
  dplyr::summarise(age = min(age)) %>% 
  dplyr::ungroup() %>% 
  unique() %>% 
  na.omit()

###########################################################
## Set up data for age matching
###########################################################

# Get tremor features from synapse and count number of windows available for each hc
ftrs.id = c(handToNose_left = 'syn12104398', handToNose_right = 'syn12104396')
all.used.ids = c(all.used.ids, as.character(ftrs.id))

hc.nc = purrr::map(ftrs.id, function(id){
  fread(synapser::synGet(id)$path, fill = T, select = c('recordId', 'window', 'axis')) %>%
    unique()
}) %>%
  data.table::rbindlist(idcol = 'Assay')

hc.nc <- hc.nc %>%
  dplyr::inner_join(tremor.tbl %>% dplyr::filter(recordId %in% hc.nc$recordId) %>%
                      dplyr::select(recordId, healthCode))
hc.nc <- hc.nc %>%
  dplyr::left_join(demo.tbl %>% dplyr::filter(healthCode %in% hc.nc$healthCode) %>%
                     dplyr::select(healthCode, gender.json.answer, dataGroups))
hc.nc <- hc.nc %>%
  dplyr::left_join(profile.tbl.cleaned %>% dplyr::filter(healthCode %in% hc.nc$healthCode)) %>%
  dplyr::rename(MS = dataGroups, gender=gender.json.answer) %>%
  dplyr::filter(!is.na(MS), !is.na(gender),
                MS %in% c('control', 'ms_patient'),
                gender %in% c('Female', 'Male')) %>%
  dplyr::count(healthCode, gender, MS,age) %>%  # Not counting across Assay (left/right) as we will consider them as repeat measures
  unique()

# High frequency filter (atleast 1 record with 3 windows in each axis, in total 12 windows in both hands)
hc.hfreq = hc.nc %>% na.omit() %>% 
  dplyr::select(healthCode, age, gender, MS, n) %>% 
  dplyr::filter(n > 0) %>% 
  unique()

###########################################################
## Age based matching (exact)
###########################################################
# dfOrig <- downloadFile('syn12118582')
dfOrig.kinetic <- read.table('hcwiseSummaryFeaturesAllAssays.tsv', sep='\t', header=TRUE) %>% 
  dplyr::filter(assay == 'kinetic') %>% 
  unique()

hc.nc.mpower <- dfOrig.kinetic %>%
  dplyr::filter(PD %in% c('control')) %>% 
  dplyr::select(healthCode, age, MS = PD, gender) %>% 
  na.omit() %>% 
  unique() %>% 
  droplevels() %>% 
  dplyr::mutate(n = 3) # I just made all the healthCodes in mpower have 3 as the number of records, infact this is higher

# Add mpower controls to elevateMS dataset before going through
# age matching

hc.hfreq.all <- rbind(hc.hfreq, hc.nc.mpower) %>% unique()

set.seed(123456789)
dat = hc.hfreq.all %>%
  dplyr::mutate(MS = factor(MS, levels = c('control','ms_patient')),
                MS = as.numeric(MS) - 1) %>%
  unique()
age.matched = matchit(MS ~ age, data = data.frame(dat), method = "exact")
dat = match.data(age.matched)

# Break ties using number of records
set.seed(123456789)
age.records.matched = plyr::ddply(dat, .(subclass), .fun = function(x){
  dat1 = x %>%
    dplyr::select(healthCode, MS, n, age, gender) %>%
    unique() 
  records.matched = matchit(MS ~ n, data = dat1)
  dat1 = match.data(records.matched)
})
p1 = ggplot(age.records.matched, aes(x = factor(MS), y = age)) + geom_boxplot()
p2 = ggplot(age.records.matched, aes(x = MS)) + geom_bar()
p3 = ggplot(age.records.matched, aes(x = log10(n), color = factor(MS))) + geom_density() + xlab('log10(# Windows)')
ggpubr::ggarrange(p2,p1,p3, ncol = 3, nrow = 1)
```

```{r get.ftrs}
### Extract features elevateMS
###########################################################
## Download and set up features for elevateMS
###########################################################
# Get retained recordID
# Select retained recordIds in elevateMS
rid.retained = tremor.tbl %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>%
  dplyr::select(recordId, healthCode)

# Load features from synapse
ftrs = purrr::map(ftrs.id, function(id){
  fread(synapser::synGet(id)$path, fill = TRUE) %>%
    dplyr::filter(IMF %in% c(1,2)) %>%
    unique()
  }) %>%
  data.table::rbindlist(idcol = 'Assay') %>%
  dplyr::inner_join(tremor.tbl %>%
                      dplyr::select(recordId, healthCode, dataGroups)) %>%
  dplyr::left_join(demo.tbl %>%
                     dplyr::select(healthCode, gender.json.answer) %>%
                     unique()) %>%
  dplyr::rename(MS = dataGroups, gender = gender.json.answer) %>% 
  dplyr::mutate(IMF = paste0('IMF',IMF)) %>% 
  dplyr::filter(MS %in% c('ms_patient','control'))

ftrs$energy.tm <- as.numeric(ftrs$energy.tm)
```

```{r fxn}
## Function block
## Function to classify case-control differences in each assay (balanced data)
rfModelCaseControl <- function(dat, responseNames, factorFeatures, contFeatures, alwaysSplitVariables){
  # Get data and format data
  dat[,c(factorFeatures, responseNames)] = lapply(dat[,c(factorFeatures, responseNames), drop = F], factor)
  dat[,contFeatures] = lapply(dat[,contFeatures, drop = F], as.numeric)
  
  # Split training and testing data (based on healthCodes)
  hc = as.character(unique(dat$healthCode))
  hc.train = sample(hc, round(length(hc)*0.6))
  hc.test = setdiff(hc, hc.train)
  
  # Get training data
  dat.train = dat %>%
    dplyr::filter(healthCode %in% hc.train) %>%
    dplyr::select(one_of(factorFeatures), one_of(contFeatures), 'MS') %>%
    dplyr::mutate(MS = as.numeric(MS) - 1)
  dat.train[is.na(dat.train)] = 0
  
  # Get testing data
  dat.test = dat %>%
    dplyr::filter(healthCode %in% hc.test) %>%
    dplyr::select(one_of(factorFeatures), one_of(contFeatures), 'MS') %>%
    dplyr::mutate(MS = as.numeric(MS) - 1)
  dat.test[is.na(dat.test)] = 0

  tryCatch({
    # Fit a rf model
    rf.mdl = ranger(dependent.variable.name = responseNames, 
                    data = dat.train, 
                    num.trees = 1000,
                    importance = 'permutation',
                    seed = 123456,
                    classification = TRUE,
                    write.forest = TRUE)
  
    # Predict reponse of test data
    rf.prediction = predict(rf.mdl, 
                            data = dat.test,
                            predict.all = T,
                            type = 'response',
                            seed = 123456)
  
    pred.probabilities = rowSums(rf.prediction$predictions == 1, na.rm = T)/rf.prediction$num.trees
  
    mdl.prediction = pred.probabilities %>%
      prediction(predictions = ., labels = dat.test[,'MS'], label.ordering = c(0,1))
    
    auroc = data.frame(metricName = 'auroc', 
                       value = performance(mdl.prediction, 'auc')@y.values[[1]])
    acc = data.frame(metricName = 'acc',
                     value = max(performance(mdl.prediction, 'acc')@y.values[[1]]))
    
    rec = performance(mdl.prediction, 'rec')@y.values[[1]]
    pre = performance(mdl.prediction, 'prec')@y.values[[1]]
    
    aupr = data.frame(metricName = 'aupr', 
                      value = caTools::trapz(rec[!is.nan(rec) & !is.nan(pre)], 
                                             pre[!is.nan(rec) & !is.nan(pre)]))
    mdl.performance = rbind(auroc, acc, aupr)
  
    mdl.coeff = ranger::importance(rf.mdl) %>%
      rownameToFirstColumn('featureNames') %>%
      dplyr::rename(value = DF)
  
    mdl.param = rbind(
      data.frame(metricName = 'nData', value = dim(dat)[1]),
      data.frame(metricName = 'nTrain', value = dim(dat.train)[1]),
      data.frame(metricName = 'nTest', value = dim(dat.test)[1]),
      data.frame(metricName = 'nTrainControl', value = sum(dat.train[,'MS'] == 0)),
      data.frame(metricName = 'nTestControl', value = sum(dat.test[,'MS'] == 1))
    )
    
    all.results = list(param = mdl.param, performance = mdl.performance, coeff = mdl.coeff)
  }, error = function(e){
    all.results = list(param = NA, performance = NA, coeff = NA)
  })
}

## Function to display number of data points
printNumData <- function(mdls){
  mapply(function(x, y){
    x[[1]]$param %>% 
      plyr::rename(c('value'=y))
  }, mdls, names(mdls), SIMPLIFY = F) %>% 
    plyr::join_all()
}

## Function to plot and display average model performance
printModelPerformance <- function(mdl, assayName){
  tmp = purrr::imap(mdl, function(x, y){
    names(x) = paste0('perm.',1:length(x))
    tmp = purrr::imap(x, function(y,z){
      
      datTempInner <- y$performance
      if(is.na(datTempInner)){
        datTempInner <- data.frame(c('auroc','acc','aupr'),rep(NA,3))
        colnames(datTempInner) <- c('metricName','value')
      }
      return(datTempInner)
    }) %>%
      plyr::join_all(by = 'metricName')}) 
  
  ##### DATA RESHAPE IS NEEDED
  tmpa <- data.frame(rep(NA,3*(length(tmp$alternateModel)-1)),rep(NA,3*(length(tmp$alternateModel)-1)),rep(NA,3*(length(tmp$alternateModel)-1)))
  colnames(tmpa) <- c('metricName','value','modelName')
  tmpa$metricName <- rep(tmp$alternateModel$metricName,length(tmp$alternateModel)-1)
  a <- unlist(tmp$alternateModel)
  tmpa$value <- a[4:(3*length(tmp$alternateModel))]
  tmpa$modelName <- rep('alternateModel', length(tmpa$value))
  
  tmpb <- data.frame(rep(NA,3*(length(tmp$alternateModel)-1)),rep(NA,3*(length(tmp$alternateModel)-1)),rep(NA,3*(length(tmp$alternateModel)-1)))
  colnames(tmpb) <- c('metricName','value','modelName')
  tmpb$metricName <- rep(tmp$nullModel$metricName,length(tmp$alternateModel)-1)
  a <- unlist(tmp$nullModel)
  tmpb$value <- a[4:(3*length(tmp$nullModel))]
  tmpb$modelName <- rep('nullModel', length(tmpb$value))
  
  tmp <- rbind(tmpa,tmpb)
  
  p = ggplot(tmp, aes(x = metricName, y = value)) + geom_boxplot(aes(color = modelName), position = position_identity())
  p = p + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ylim(0,1) + ggtitle(assayName) + scale_color_manual(values = c('black','grey60'))
  
  tmp = dplyr::group_by(tmp, modelName, metricName) %>%
    dplyr::summarise(value = round(mean(value, na.rm = T),4)) %>%
    tidyr::spread(modelName, value) 
  return(list(plot = p, performance = tmp))
}

## Function to print top n features 
printTopFeatures <- function(mdl){
  purrr::imap(mdl, function(x, i){
    x$coeff %>%
      plyr::rename(c('value' = i))
    }) %>%
    plyr::join_all(by = 'featureNames') %>%
    tidyr::gather(iter, value, -featureNames) %>%
    dplyr::group_by(featureNames) %>%
    dplyr::summarise(value = sum(value != 0, na.rm = T)/100) %>%
    dplyr::arrange(desc(value)) %>%
    dplyr::filter(value >= 0.9) %>%
    dplyr::select(featureNames) %>%
    unlist %>% 
    paste(collapse = ', ')
}

## Get top n percent of features 
getTopNPercentOfFtrs <- function(mdl, n = 70){
  purrr::imap(mdl, function(x, i){
    x$coeff %>%
      plyr::rename(c('value' = i))
  }) %>%
    plyr::join_all() %>%
    tidyr::gather(iter, value, -featureNames) %>%
    dplyr::group_by(featureNames) %>%
    dplyr::summarise(value = sum(value != 0, na.rm = T)/100) %>%
    dplyr::arrange(desc(value)) %>%
    dplyr::filter(value >= n/100) %>%
    dplyr::select(featureNames) %>%
    unlist %>% 
    setdiff('(Intercept)')
}
```

### Features considered for this analysis
```{r ftr.def, echo = TRUE}
# Split features based componenet
ftrs.names = c("mean.tm", "median.tm", "mode.tm", "mx.tm", "mn.tm", "sd.tm", "skewness.tm", 
               "kurtosis.tm", "Q25.tm", "Q75.tm", "range.tm", "rough.tm", "rugo.tm", "energy.tm",
               "mobility.tm", "mtkeo.tm", "dfa.tm", "rmsmag.tm", "IQR.tm", "complexity.tm",
               "mn.fr", "mx.fr", "sd.fr", "sem.fr", "md.fr", "mod.fr", "Q25.fr", "Q75.fr", 
               "IQR.fr", "cent.fr", "skew.fr", "kurt.fr", "sfm.fr", "sh.fr", "ewt.permEnt.fr",
               "ewt.shannonEnt.fr", "ewt.simpsonEnt.fr", "ewt.renyiEnt.fr", "ewt.tsallisEnt.fr", 
               "EnergyInBand1", "EnergyInBand1_5", "EnergyInBand2", "EnergyInBand2_5", "EnergyInBand3",
               "EnergyInBand3_5", "EnergyInBand4", "EnergyInBand4_5", "EnergyInBand5", "EnergyInBand5_5",
               "EnergyInBand6", "EnergyInBand6_5", "EnergyInBand7", "EnergyInBand7_5", "EnergyInBand8",
               "EnergyInBand8_5", "EnergyInBand9", "EnergyInBand9_5", "EnergyInBand10", "EnergyInBand10_5",
               "EnergyInBand11", "EnergyInBand11_5", "EnergyInBand12", "EnergyInBand12_5", "EnergyInBand13",
               "EnergyInBand13_5", "EnergyInBand14", "EnergyInBand14_5", "EnergyInBand15", "EnergyInBand15_5",
               "EnergyInBand16", "EnergyInBand16_5", "EnergyInBand17", "EnergyInBand17_5", "EnergyInBand18",
               "EnergyInBand18_5", "EnergyInBand19", "EnergyInBand19_5", "EnergyInBand20", "EnergyInBand20_5",
               "EnergyInBand21", "EnergyInBand21_5", "EnergyInBand22", "EnergyInBand22_5", "EnergyInBand23",
               "EnergyInBand23_5", "EnergyInBand24", "EnergyInBand24_5")
```

#### Combine energy features in to bands of 1 Hz before feature selection for elevateMS
```{r all.energy}
###########################################################
## Combine energy features in to 1 Hz band
###########################################################
energy.ftr = ftrs %>%
  tidyr::unite(rid, recordId, Assay, sensor, measurementType, IMF, axis, window, sep = '.') %>%
  dplyr::select(rid, contains('EnergyInBand'))

energy.ftr.cmbn = purrr::map2(seq(1,24,by=1),
                              seq(1.5,24.5,by=1),
                              function(x, y, innerFtr){
                                ind = paste0('EnergyInBand',gsub('\\.','_', seq(x, y, by=0.5)))
                                innerFtr = innerFtr %>%
                                  dplyr::select(rid, one_of(ind))
                                rowSums(innerFtr[,2:3], na.rm = T)
                              }, energy.ftr) %>%
  do.call(cbind,.)
colnames(energy.ftr.cmbn) = paste0('EnergyInBand', seq(1,24,by=1))
energy.ftr.cmbn = cbind(data.frame(rid = energy.ftr$rid), energy.ftr.cmbn)
```

#### Find median and iqr of each feature (per healthCode) for elevateMS
```{r ftr.summarize}
###########################################################
## Summarize features for elevateMS
###########################################################
# Get kinetic tremor features
kinetic.ftr = ftrs %>%
  tidyr::unite(rid, recordId, Assay, sensor, measurementType, IMF, axis, window, sep = '.') %>%
  dplyr::select(-contains('EnergyInBand')) %>%
  dplyr::left_join(energy.ftr.cmbn) %>%
  tidyr::separate(rid, c('recordId', 'Assay', 'sensor', 'measurementType', 'IMF', 'axis', 'window'), sep = '\\.') %>%
  dplyr::select(-recordId, -Assay, -axis, -window) %>%
  tidyr::gather(Feature, Value, -healthCode, -gender, -MS, -sensor, -measurementType, -IMF) %>%
  dplyr::group_by(Feature, healthCode, gender, MS, sensor, measurementType, IMF) %>%
  dplyr::summarise(iqr = stats::IQR(Value, na.rm = T),
                   md = stats::median(Value, na.rm = T))

# Get kinetic data and covariates seperately
kinetic.cov = kinetic.ftr %>%
  dplyr::ungroup() %>%
  dplyr::select(healthCode, gender, MS) %>%
  unique() %>%
  purrr::map_df(factor) %>%
  as.data.frame()
rownames(kinetic.cov) = kinetic.cov$healthCode

# Get median of features
kinetic.ftr.md = kinetic.ftr %>%
  dplyr::ungroup() %>%
  dplyr::select(healthCode, sensor, measurementType, Feature, IMF, md) %>%
  dplyr::mutate(type = 'md') %>%
  tidyr::unite(nFeature, Feature, IMF, type, sep = '.') %>%
  tidyr::spread(nFeature, md)

# Get iqr of features
kinetic.ftr.iqr = kinetic.ftr %>%
  dplyr::ungroup() %>%
  dplyr::select(healthCode, sensor, measurementType, Feature, IMF, iqr) %>%
  dplyr::mutate(type = 'iqr') %>%
  tidyr::unite(nFeature, Feature, IMF, type, sep = '.') %>%
  tidyr::spread(nFeature, iqr)

# Combine median and iqr features (all data)
kinetic.ftr = dplyr::inner_join(kinetic.ftr.md, kinetic.ftr.iqr)

# Remove linearly associated features
tmp.mat = kinetic.ftr %>%
  dplyr::select(-healthCode, -sensor, -measurementType)
lm.combo = caret::findLinearCombos(tmp.mat)

kinetic.ftr.all = kinetic.ftr %>%
  dplyr::select(-one_of(colnames(tmp.mat)[lm.combo$remove])) %>%
  tidyr::gather(Feature, Value, -healthCode, -sensor, -measurementType) %>%
  tidyr::unite(featureName, Feature, measurementType, sensor, sep = '_') %>%
  tidyr::spread(featureName, Value)
 
# kinetic.ftr <- kinetic.ftr.all %>% dplyr::filter(healthCode %in% age.records.matched$healthCode)
```


```{r filter.healthcodes.mpower}
## Get and format data for mPower
### Age match case and control healthCodes
###########################################################
## Download and set up features for mPower
###########################################################

############### Reshape Data
# dfOrig <- downloadFile('syn12118582')
dfOrig <- read.table('hcwiseSummaryFeaturesAllAssays.tsv', sep='\t', header=TRUE)

colsToNotSelect <- c('assay','PD','age','gender')
dfCommonCols <- dfOrig %>% dplyr::select(c('healthCode',colsToNotSelect))

df <- dfOrig %>% 
  dplyr::select(setdiff(colnames(dfOrig), colsToNotSelect)) %>% 
  dplyr::filter(medTimepoint == 'preMed')
dfKinetic <- dfOrig %>% 
  dplyr::filter(assay == 'kinetic', 
                medTimepoint == 'preMed') %>%
  dplyr::select(c('healthCode',colnames(df)[grep('IMF',colnames(df))]))

tremor.tbl.mpower.syn <- synapser::synTableQuery(paste0("SELECT * FROM ", 'syn10676309'))
tremor.tbl.mpower <- tremor.tbl.mpower.syn$asDataFrame() %>%
  dplyr::select(recordId, healthCode, momentInDayFormat.json.choiceAnswers) %>%
  dplyr::rename(medTimepoint = momentInDayFormat.json.choiceAnswers) %>%
  dplyr::filter(medTimepoint != "Just after taking Parkinson medication (at your best)") %>%
  unique()

load('AllRanking.RData') #syn12038526
nFeatures = 100 # max number of features to choose per assay

kineticFeaturesIMF1 <- ranka[,3] %>% na.omit() 
kineticFeaturesIMF2 <- ranka[,4] %>% na.omit() 
maxRank <- max(c(kineticFeaturesIMF1,kineticFeaturesIMF2))
kineticFeaturesIMF1 <- kineticFeaturesIMF1[kineticFeaturesIMF1 > (maxRank-nFeatures)] %>% names()
kineticFeaturesIMF2 <- kineticFeaturesIMF2[kineticFeaturesIMF2 > (maxRank-nFeatures)] %>% names()

kineticFeaturesToSelect <- c()
for (ftr in colnames(dfKinetic)){
  ftr1 <- gsub('IMF1.','',ftr)
  # print(ftr1)
  if(ftr1 %in% kineticFeaturesIMF1){
    kineticFeaturesToSelect <- c(kineticFeaturesToSelect,ftr)
  }
  ftr1 <- gsub('IMF2.','',ftr)
  # print(ftr1)
  if(ftr1 %in% kineticFeaturesIMF2){
    kineticFeaturesToSelect <- c(kineticFeaturesToSelect,ftr)
  }
}

dfKinetic <- dfKinetic %>% dplyr::select(c('healthCode',kineticFeaturesToSelect))
dfKinetic$healthCode <- droplevels(dfKinetic$healthCode)
dfKinetic.orig <- dfKinetic

colnames(dfKinetic) <- paste0(colnames(dfKinetic),'.kinetic')
colnames(dfKinetic)[1] <- 'healthCode'

retained.hcs <- c(age.records.matched$healthCode, tremor.tbl$healthCode) %>% 
  unique()

# For training data add both the ftrs from elevateMS and the controls from 
# mpower and filter to get the whole age matched data
dfMainTrain <- dplyr::full_join(kinetic.ftr.all, dfKinetic.orig) %>% 
  dplyr::filter(healthCode %in% retained.hcs) %>%
  dplyr::select(c('healthCode',kineticFeaturesToSelect)) %>% 
  unique()

```
#### Compute model performance using identified features (from mpoweranalysis)
##### Evaluate model on training set (mpower + elevateMS) (alternate vs null model)
```{r all.rf.model}
nperm = 100
set.seed(123456789)
exportPackages = c('dplyr','plyr','data.table','CovariateAnalysis', 'ranger', 'ROCR')
ftrs.to.evaluate = kineticFeaturesToSelect %>% unique()

# Get training data
dat.train = dfMainTrain %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
  dplyr::select(one_of(ftrs.to.evaluate)) %>%
  apply(2, scale) 
dat.train = cbind(data.frame(healthCode = dfMainTrain %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
    dplyr::select(healthCode)), dat.train) %>%
  dplyr::left_join(hc.hfreq.all %>% 
                     dplyr::filter(healthCode %in% dfMainTrain$healthCode)) %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
  as.data.frame() %>% droplevels() %>% 
  dplyr::mutate(MS = factor(MS, levels = c('control','ms_patient')),
                MS = as.numeric(MS) - 1) %>% dplyr::select(-age, -n) %>% 
  na.omit() %>%
  unique()

# Run random forest model
rf.mdl.caret =  c('alternateModel','nullModel') %>%
  plyr::llply(function(mdl.tag, datTrain, nperm, ftrsNames, rfModelCaseControl){
    mdl = plyr::llply(seq(nperm), .fun = function(iperm, mdlTag, datTrain, ftrsNames, rfModelCaseControl){
      # Randomise case control label per healthCode
      if(mdlTag == 'nullModel')
        datTrain$MS = sample(datTrain$MS) 
      
      rfModelCaseControl(datTrain,  
                         responseNames = 'MS',
                         factorFeatures = c('gender'),
                         contFeatures = ftrsNames)
    }, mdl.tag, datTrain, ftrsNames, rfModelCaseControl,
    .parallel = F, 
    .paropts = list(.packages = exportPackages, 
                    .export = c('rfModelCaseControl')))
    return(mdl)
  },
  dat.train, nperm, setdiff(ftrs.to.evaluate, c('age', 'gender')), rfModelCaseControl,
  .paropts = list(.packages = exportPackages, 
                  .export = c('rfModelCaseControl')))
names(rf.mdl.caret) = c('alternateModel','nullModel')
```

##### Model performance
```{r all.print.results, fig.height=5, fig.width=12, cache = FALSE}
# Display number of data points used
writeLines('Number of data points used')
kable(rf.mdl.caret$alternateModel[[1]]$param)

# Plot auroc and aupr
writeLines('Average performance on 100 runs')
tmp = printModelPerformance(rf.mdl.caret, 'Caret based features')
p = gridExtra::tableGrob(tmp$performance, rows = NULL)
p = ggpubr::ggarrange(tmp$plot, p, nrow = 1, ncol = 2)
p

# Plot important variables in all model
writeLines('Features present in 70% of the models are')
printTopFeatures(rf.mdl.caret$alternateModel) %>%
  paste(collapse = ', ') %>%
  data.frame(Important_Features = .) %>%
  kable()
```
#### Evaluate Kinetic Model on elevateMS by training on mPower+elevateMS age matched data
```{r model.result}
nperm = 100
set.seed(123456789)
exportPackages = c('dplyr','plyr','data.table','CovariateAnalysis', 'ranger', 'ROCR')

# ftrsToSelect <- paste0(kineticFeaturesToSelect,'.kinetic')
ftrsToSelect <- kineticFeaturesToSelect
  
# Get training data
dat.train = dfMainTrain %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
  dplyr::select(one_of(ftrsToSelect)) %>%
  apply(2, scale) 
dat.train = cbind(data.frame(healthCode = dfMainTrain %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
    dplyr::select(healthCode)), dat.train) %>%
  dplyr::left_join(hc.hfreq.all %>% 
                     dplyr::filter(healthCode %in% dfMainTrain$healthCode)) %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
  as.data.frame() %>% droplevels() %>% 
  dplyr::mutate(MS = factor(MS, levels = c('control','ms_patient')),
                MS = as.numeric(MS) - 1) %>% dplyr::select(-age, -n) %>% 
  na.omit() %>%
  unique()

# colnames(dat.train) <- gsub('.kinetic','',colnames(dat.train))
dat.test <- dfMainTrain %>%
  dplyr::filter(!(healthCode %in% age.records.matched$healthCode)) %>% 
    dplyr::select(one_of(kineticFeaturesToSelect)) %>%
      apply(2, scale)
dat.test = cbind(data.frame(healthCode = dfMainTrain %>%
  dplyr::filter(!(healthCode %in% age.records.matched$healthCode)) %>% 
    dplyr::select(healthCode)), dat.test) %>%
   dplyr::left_join(kinetic.cov) %>%
  as.data.frame() %>% droplevels() %>%
  dplyr::mutate(MS = factor(MS, levels = c('control','ms_patient')),
                MS = as.numeric(MS) - 1) %>% 
  dplyr::filter(healthCode %in% tremor.tbl$healthCode) %>% 
  dplyr::filter(!(healthCode %in% age.records.matched$healthCode)) 
# %>%
#    na.omit()

dat.test$gender[is.na(dat.test$gender)] <- 'Male'
# We have a lot of missing gender data, I'm just defaulting it to Male
# It does not have any effect on the prediction if I choose all to be
# Male or Female, I have tried both

    rf.mdl = ranger(dependent.variable.name = 'MS', 
                    data = dat.train, 
                    num.trees = 1000,
                    importance = 'permutation',
                    seed = 123456,
                    classification = TRUE,
                    write.forest = TRUE)
  
    # Predict reponse of test data
    rf.prediction = predict(rf.mdl, 
                            data = dat.test,
                            predict.all = T,
                            type = 'response',
                            seed = 123456)
        pred.probabilities = rowSums(rf.prediction$predictions == 1, na.rm = T)/rf.prediction$num.trees
  
    mdl.prediction = pred.probabilities %>%
      prediction(predictions = ., labels = dat.test[,'MS'], label.ordering = c(0,1))
    
    auroc = data.frame(metricName = 'auroc', 
                       value = performance(mdl.prediction, 'auc')@y.values[[1]])
    acc = data.frame(metricName = 'acc',
                     value = max(performance(mdl.prediction, 'acc')@y.values[[1]]))
    
    rec = performance(mdl.prediction, 'rec')@y.values[[1]]
    pre = performance(mdl.prediction, 'prec')@y.values[[1]]
    
    aupr = data.frame(metricName = 'aupr', 
                      value = caTools::trapz(rec[!is.nan(rec) & !is.nan(pre)], 
                                             pre[!is.nan(rec) & !is.nan(pre)]))
    mdl.performance = rbind(auroc, acc, aupr)
    
writeLines(c('Number of Kinetic features chosen for analysis : ',as.character(length(kineticFeaturesToSelect))))
writeLines(c('Number of healthCodes in testset : ',as.character(length(dat.test$healthCode))))
writeLines('Prediction on elevateMS after training on age matched mPower + elevateMS data')
kable(mdl.performance)
```

#### Evaluate Kinetic Model on elevateMS by training on mPower+elevateMS age matched data (removing gender == NA data)
```{r model.result.gender}
nperm = 100
set.seed(123456789)
exportPackages = c('dplyr','plyr','data.table','CovariateAnalysis', 'ranger', 'ROCR')

# ftrsToSelect <- paste0(kineticFeaturesToSelect,'.kinetic')
ftrsToSelect <- kineticFeaturesToSelect
  
# Get training data
dat.train = dfMainTrain %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
  dplyr::select(one_of(ftrsToSelect)) %>%
  apply(2, scale) 
dat.train = cbind(data.frame(healthCode = dfMainTrain %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
    dplyr::select(healthCode)), dat.train) %>%
  dplyr::left_join(hc.hfreq.all %>% 
                     dplyr::filter(healthCode %in% dfMainTrain$healthCode)) %>%
  dplyr::filter(healthCode %in% age.records.matched$healthCode) %>% 
  as.data.frame() %>% droplevels() %>% 
  dplyr::mutate(MS = factor(MS, levels = c('control','ms_patient')),
                MS = as.numeric(MS) - 1) %>% dplyr::select(-age, -n) %>% 
  na.omit() %>%
  unique()

# colnames(dat.train) <- gsub('.kinetic','',colnames(dat.train))
dat.test <- dfMainTrain %>%
  dplyr::filter(!(healthCode %in% age.records.matched$healthCode)) %>% 
  dplyr::left_join(kinetic.cov %>% dplyr::select(healthCode, gender)) %>% 
  na.omit() %>% 
    dplyr::select(one_of(kineticFeaturesToSelect)) %>%
      apply(2, scale)
dat.test = cbind(data.frame(healthCode = dfMainTrain %>%
  dplyr::filter(!(healthCode %in% age.records.matched$healthCode)) %>% 
      dplyr::left_join(kinetic.cov %>% dplyr::select(healthCode, gender)) %>% 
    na.omit() %>% 
    dplyr::select(healthCode)), dat.test) %>%
   dplyr::left_join(kinetic.cov) %>%
  as.data.frame() %>% droplevels() %>%
  dplyr::mutate(MS = factor(MS, levels = c('control','ms_patient')),
                MS = as.numeric(MS) - 1) %>% 
  dplyr::filter(healthCode %in% tremor.tbl$healthCode) %>% 
  dplyr::filter(!(healthCode %in% age.records.matched$healthCode)) %>%
   na.omit()

    rf.mdl = ranger(dependent.variable.name = 'MS', 
                    data = dat.train, 
                    num.trees = 1000,
                    importance = 'permutation',
                    seed = 123456,
                    classification = TRUE,
                    write.forest = TRUE)
  
    # Predict reponse of test data
    rf.prediction = predict(rf.mdl, 
                            data = dat.test,
                            predict.all = T,
                            type = 'response',
                            seed = 123456)
        pred.probabilities = rowSums(rf.prediction$predictions == 1, na.rm = T)/rf.prediction$num.trees
  
    mdl.prediction = pred.probabilities %>%
      prediction(predictions = ., labels = dat.test[,'MS'], label.ordering = c(0,1))
    
    auroc = data.frame(metricName = 'auroc', 
                       value = performance(mdl.prediction, 'auc')@y.values[[1]])
    acc = data.frame(metricName = 'acc',
                     value = max(performance(mdl.prediction, 'acc')@y.values[[1]]))
    
    rec = performance(mdl.prediction, 'rec')@y.values[[1]]
    pre = performance(mdl.prediction, 'prec')@y.values[[1]]
    
    aupr = data.frame(metricName = 'aupr', 
                      value = caTools::trapz(rec[!is.nan(rec) & !is.nan(pre)], 
                                             pre[!is.nan(rec) & !is.nan(pre)]))
    mdl.performance = rbind(auroc, acc, aupr)
    
writeLines(c('Number of Kinetic features chosen for analysis : ',as.character(length(kineticFeaturesToSelect))))
writeLines(c('Number of healthCodes in testset : ',as.character(length(dat.test$healthCode))))
writeLines('Prediction on elevateMS after training on age matched mPower + elevateMS data')
kable(mdl.performance)
```

### Store results in synapse
```{r syn.store, cache=FALSE, eval = TRUE}
CODE = Folder(name = 'elevateMS action tremor test (mpower PD action tremor train)', 
              parentId = "syn11315047")
CODE = synStore(CODE)

# Save all coefficients and AUCs as RData
save(list = c('rf.mdl'), 
     file = 'AllImportanceScore.RData')
obj = File('AllImportanceScore.RData', 
           name = 'All importance scores with AUC', 
           parentId = CODE$properties$id)
obj = synStore(obj,activityName = activityName, 
               activityDescription = activityDescription, 
               used = all.used.ids, executed = thisFile)
res = paste0('https://www.synapse.org/#!Synapse:',obj$properties$id)
# stopCluster(cl)
```
[Results](`r res`)

### Source code in github
[Source Code](`r thisFile`)